---
title: "esm_manuscript_revision2"
author: "Katy Dynarski"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(here)
library(janitor)
library(paletteer)
library(aqp)
library(broom)
library(broom.mixed)
library(lme4)
library(lmerTest)
library(multcomp)
library(ggeffects)
library(ggpubr)
library(sf)
library(ggspatial)
library(ggeasy)
library(ggrepel)
library(flextable)
library(patchwork)
library(multcomp)
library(multcompView)
library(PNWColors)
library(tidyverse)

# Data
# Vector for method order for consistent plotting
method_factor_order <- c("esm2_min_project", "esm2_min_treatment", "esm2_mean_project", "esm2_mean_treatment", "fd_fd_fd")

# Vector for project order for consistent plotting
project_plotting_order <- c("UConn", "Illinois", "KansasState", "UnivOfMinnesota", "NCState", 
                            "OregonStateJory", "OregonStateWoodburn","TexasA&MPt-1", "WashingtonState")

esm <- read.csv(here("data_processed", "esm_standard_depths_revision.csv")) %>%
  filter(project!="UTRGV") %>% # don't include UTRGV for this analysis - bulk density is from SSURGO
  filter(method!="esm1") %>% # not including ESM classic in this analysis - just compare cubic spline
  unite("method_long", c("method", "ref_stat"), sep="_", remove = FALSE) %>%
  unite("method_longest", c("method", "ref_stat", "ref_data"), sep="_", remove = FALSE) %>%
  mutate(depth_std = case_when(layer==1 ~ 5,
                               layer==2 ~ 10,
                               layer==3 ~ 30,
                               layer==4 ~ 60),
         esm_depth = factor(esm_depth, levels=c("0-5 cm", "5-10 cm", "10-30 cm", "30-60 cm")),
         label = factor(label, levels=c("BAU", "SHM", "Ref")),
         project = factor(project, levels=project_plotting_order),
         method_longest = factor(method_longest, levels=method_factor_order))
  

horizon_data <- read.csv(here("data_processed", "soc_horizon_esm_input.csv"))
project <- read.csv(here("data_processed", "project_data_meta.csv"))

# List of projects
projects <- esm %>% 
  distinct(project) %>% 
  mutate(project = as.character(project)) %>%
  pull()

inc_depth_labels <- c(
  "5" = "ESM Increment\nDepth: 0-5 cm",
  "10" = "ESM Increment\nDepth: 5-10 cm",
  "30" = "ESM Increment\nDepth: 10-30 cm",
  "60" = "ESM Increment\nDepth: 30-60 cm"
)

cum_depth_labels <- c(
  "5" = "ESM Cumulative\nDepth: 0-5 cm",
  "10" = "ESM Cumulative\nDepth: 0-10 cm",
  "30" = "ESM Cumulative\nDepth: 0-30 cm",
  "60" = "ESM Cumulative\nDepth: 0-60 cm"
)
  
method_labels <- c(
  "esm2_min_project" = expression("ESM"["project, min"]),      
  "esm2_mean_project" = expression("ESM"["project, mean"]), 
  "esm2_min_treatment"= expression("ESM"["control, min"]),   
  "esm2_mean_treatment" = expression("ESM"["control, mean"]),
  "fd_fd_fd" = "Fixed depth"
)

method_labels2 <- c(
  "esm2_min_project" = "ESM(project, min)",      
  "esm2_mean_project" = "ESM(project, mean)", 
  "esm2_min_treatment"= "ESM(control, min)",   
  "esm2_mean_treatment" = "ESM(control, mean)",
  "fd_fd_fd" = "Fixed depth"
)

project_labels_esm <- c("UnivOfMinnesota" = "U. Minnesota",
                    "WashingtonState" = "Washington State",
                    "TexasA&MPt-1" = "Texas A&M",
                    "NCState" = "North Carolina State",
                    "Illinois" = "U. Illinois",
                    "OregonStateJory" = "Oregon State - Jory",
                    "OregonStateWoodburn" = "Oregon State - Woodburn",
                    "KansasState" = "Kansas State",
                    "UConn" = "U. Connecticut")

project_labels_esm_df <- data.frame(project_labels_esm) %>%
  rownames_to_column()

mgmt_pal <- c("#FED789FF","#72874EFF","#476F84FF")

theme_katy <- function(base_size=8) {
  theme_classic(base_size=base_size) %+replace%
    theme(# Legend
      legend.title=element_text(size=rel(1)), legend.text=element_text(size=rel(.8)),
      #axes                      
      axis.text=element_text(size=rel(.8)),
      axis.title=element_text(size=rel(1)),
      axis.line=element_line(linewidth=0.25),
      axis.ticks=element_line(linewidth=0.25),
      strip.background=element_rect(linewidth=0.3),)
}
```

# Overview

## Calculation method comparison

I calculated SOC stocks in 0-5 cm, 5-10 cm, 10-30 cm, and 30-60 cm depth increments via fixed depth and ESM methods. For the ESM calculations, I tested four different reference mass options:

* Minimum soil mass in an individual DSP4SH project (ESM project, min)
* Mean soil mass in an individual DSP4SH project (ESM project, mean)
* Minimum soil mass in the reference treatment in an individual DSP4SH project (ESM treatment, min)
* Mean soil mass in the reference treatment in an individual DSP4SH project (ESM treatment, mean)

I wanted to test just using data from the reference treatment to calculate the reference mass because a number of studies calculated their reference mass as either the mean or minimum soil mass in the treatment expected to have the lowest bulk density.

I also re-calculated bulk density after reading through the SimpleESM documentation again. We did not originally account for coarse fragment content, so I re-calculated bulk density to correct for coarse fragment content. It looks like this resolved some of the weird numbers we were getting for soils high in coarse frags (e.g. the Oregon State project).

# Climate and soils data for each project 

```{r climate means for each project}
# need to rename OSU projects in "project" dataframe and calculate mean MAT/MAP for all projects
project_renamed <- project %>%
  mutate(project = case_when(
    project=="OregonState" & soil=="Jory" ~ "OregonStateJory",
    project=="OregonState" & soil=="Woodburn" ~ "OregonStateWoodburn", 
    .default = project))

project_avg_climate <- project_renamed %>%
  group_by(project) %>%
  dplyr::summarize(avg_mat = round(mean(mat, na.rm=TRUE), 1),
         avg_ppt = round(mean(map, na.rm=TRUE), 0)) %>%
  arrange(factor(project, levels=project_plotting_order))

# write_csv(project_avg_climate, here("figs", "revision_figs", "tablesupp1_site_clim.csv"))
```

## Make map
```{r map}
# calculate average lat and long for each project
proj_loc <- project_renamed |> 
  filter(project!="UTRGV", project !="TexasA&MPt-2") |> 
  group_by(project) |> 
  summarize(mean_x = mean(pedon_x), 
            mean_y = mean(pedon_y)) |> 
  left_join(project_labels_esm_df, by=c("project" = "rowname"))

# Download map of USA from maps() and convert to sf
usa <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))

# Make map
ggplot(data=usa) +
  geom_sf(fill=NA) +
  coord_sf(xlim=c(-125.0, -66.93457), ylim=c(23.5, 49.384358)) + # set bounding box around CONUS
  annotation_north_arrow(location="bl", which_north="true", height=unit(.25, "in"), width=unit(.25, "in"),
                         pad_x = unit(0.4, "in"), pad_y = unit(0.25, "in"), style=north_arrow_fancy_orienteering) + # add north arrow
  annotation_scale(location = "bl") +
  geom_point(data=proj_loc, aes(x=mean_x, y=mean_y, color=project)) +
  geom_label_repel(data=proj_loc, aes(x=mean_x, y=mean_y, label=project_labels_esm),
                   min.segment.length = 0, seed = 42, box.padding = 0.5) +
  scale_color_manual(values=pnw_palette("Sunset2", 9)) +
  theme_katy() +
  easy_remove_axes() +
  theme(legend.position="none")
# ggsave(here("figs", "revision2_figs", "figsupp1_map.png"), width=180, height=130, units = "mm")
```

# Number of pedons and sites

```{r how many pedons and sites}
# how many projects
n_distinct(esm$project)

# how many pedons
n_distinct(esm$dsp_pedon_id)

# how many sites
# have to use some regex to extract the site name from the pedon ID since I'm missing the needed column
sites <- esm %>%
  select(project, label, dsp_pedon_id) %>%
  distinct(project, label, dsp_pedon_id) %>%
  mutate(dsp_site_id = str_remove(dsp_pedon_id, "-\\d*\\w*$")) %>%
  group_by(project, label, dsp_site_id)

# how many pedons per site?
sites %>%
  dplyr::summarize(n_pedons_per_site = n())

# how many total sites  
n_distinct(sites$dsp_site_id)

# how many total pedons
n_distinct(sites$dsp_pedon_id)

# how many total pedons per treatment
esm %>%
  select(project, label, dsp_pedon_id) %>%
  distinct(project, label, dsp_pedon_id) %>%
  group_by(label) %>%
  dplyr::summarize(n_pedons_per_treatment = n())

# table for supplement with number of sites and pedons per treatment in each project
site_table <- sites %>%
  distinct(project, label, dsp_site_id) %>%
  group_by(project, label) %>%
  dplyr::summarize(n_sites_per_treatment = n())

pedon_table <- sites %>%
  ungroup() %>%
  group_by(project, label) %>%
  dplyr::summarize(n_pedons_per_trt = n())

site_pedon_table <- site_table %>%
  left_join(pedon_table, by=c("project", "label")) %>%
  arrange(factor(project, levels=project_plotting_order), label)

# write_csv(site_pedon_table, here("figs", "revision_figs", "tablesupp2_site_pedon_numbers.csv"))
```

# Range in bulk density and reference masses

Management can influence soil bulk density, which results in different masses of soil within the same depth increment between treatments. We can check this by plotting soil bulk density under the different management treatments in each project.

Plot bulk density:

```{r calculate bulk density by depth increment, include=FALSE}
bd <- horizon_data %>%
  filter(!project=="UTRGV") %>%
  select(project, label, dsp_pedon_id, dsp_sample_id, hrzdep_t, hrzdep_b, bd_hybrid)


ggplot(bd |> filter(project=="NCState"), aes(x=bd_hybrid, fill=label)) +
  geom_density(alpha=0.7)


ggplot(bd |> filter(project=="NCState"), aes(x=hrzdep_b, y=bd_hybrid, color=label)) +
  geom_point()

# promote to SPC and dice into 1-cm increments
spc <- bd
aqp::depths(spc) <- dsp_pedon_id ~ hrzdep_t + hrzdep_b

dice <- aqp::dice(spc, fm=0:99 ~ bd_hybrid + project + label)
dice_bd <- horizons(dice)

bd_depth <- c(5, 10, 30, 60, 100)

# Initialize output vector
out <- vector("list", length(bd_depth))

for (i in seq_along(bd_depth)) {
    if (i == 1) {
      out[[i]] <- dice_bd %>%
        filter(hrzdep_t < bd_depth[[i]]) %>%
        group_by(dsp_pedon_id) %>%
        dplyr::summarize(depth_wt_bd = mean(bd_hybrid)) %>%
        mutate(depth_cat = bd_depth[[i]])
      
    } else {
      out[[i]] <- dice_bd %>%
        filter(hrzdep_t >=bd_depth[[i-1]] & hrzdep_t < bd_depth[[i]]) %>%
        group_by(dsp_pedon_id) %>%
        dplyr::summarize(depth_wt_bd = mean(bd_hybrid)) %>%
        mutate(depth_cat = bd_depth[[i]])
    }
}

out_bind <- dplyr::bind_rows(out)

bd_depth <- out_bind %>%
  left_join(select(project_renamed, project, label, dsp_pedon_id), by="dsp_pedon_id") %>%
  rename(layer = depth_cat) %>%
  mutate(depth = factor(case_when(layer == "5" ~ "0-5 cm",
                           layer == "10" ~ "5-10 cm",
                           layer == "30" ~ "10-30 cm",
                           layer == "60" ~ "30-60 cm",
                           layer == "100" ~ "60-100 cm"),
                        levels=c("0-5 cm", "5-10 cm", "10-30 cm", "30-60 cm", "60-100 cm")),
         label = factor(label, levels=c("BAU", "SHM", "Ref")),
         project = factor(project, levels=project_plotting_order))

mean_bd_depth <- bd_depth %>%
  group_by(project, label, depth) %>%
  dplyr::summarize(bd_mean = round(mean(depth_wt_bd, na.rm=TRUE), 2))
```

```{r fig 2 plot bulk density by depth increment}
# Plot bulk density for each project/treatment
ggplot(bd_depth, aes(x=depth, y=depth_wt_bd, fill=label)) +
  geom_boxplot(fatten=1.6, lwd=0.25, outlier.size=0.6) +
  facet_wrap(~project, scales="free_y", labeller=labeller(project=project_labels_esm)) +
  scale_fill_manual(values=mgmt_pal,
                     breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  labs(x="Depth", y=expression("Bulk density"~(g ~ cm^-3))) +
  theme_katy() +
  theme(axis.text.x=element_text(hjust=1, angle=45))
# ggsave(here("figs", "revision_figs", "figsupp2_bulk_density.png"), width=140, height=100, units="mm", dpi=400)
```

```{r min max bulk density for text}
bd_min_max <- bd_depth %>%
  dplyr::summarize(min = min(depth_wt_bd, na.rm=TRUE),
            max = max(depth_wt_bd, na.rm=TRUE))

# or min max mean bulk density
mean_bd_depth <- bd_depth |> 
  group_by(project, label, depth) |> 
  dplyr::summarize(mean = mean(depth_wt_bd, na.rm=TRUE))

mean_bd_depth |> 
  ungroup() |> 
  dplyr::summarize(min_mean = min(mean, na.rm=TRUE),
                   max_mean = max(mean, na.rm=TRUE))
```

## Range of reference masses for different soils:

Full table of reference masses for supplement:
```{r table s3 full ref mass table}
ref_mass_table <- esm |> 
  filter(method_longest != "fd_fd_fd") |> 
  select(project, esm_depth, method_longest, soil_mass) |> 
  distinct() |> 
  mutate(soil_mass = round(soil_mass, 0)) |> 
  pivot_wider(names_from = method_longest,
              values_from = soil_mass) |> 
  arrange(project) |> 
  relocate(project, esm_depth, esm2_min_project, esm2_min_treatment, esm2_mean_project, esm2_mean_treatment)

# write_csv(ref_mass_table, here("figs", "revision2_figs", "ref_masses_table.csv"))
```

Summary table for text:

```{r summary table of reference masses for text}
# min and max for each depth
esm |> 
  group_by(depth_std) |> 
  dplyr::summarize(min_mass = round(min(soil_mass), 0),
            max_mass = round(max(soil_mass), 0)) |> 
  flextable()

# mean reference mass for each method/depth
esm |> 
  filter(method_longest != "fd_fd_fd") |> 
  group_by(depth_std, method_longest) |> 
  dplyr::summarize(mean_mass = round(mean(soil_mass, na.rm=TRUE), 0)) |> 
  pivot_wider(names_from = method_longest, values_from = mean_mass) |> 
  flextable()
```

# How does ESM choice influence total calculated SOC stocks?

## Effect of stock calculation method on cumulative SOC stocks

What is the mean SOC stock calculated via each method?

What is the difference between the smallest and largest SOC stocks calculated for a particular pedon using these different methods?

```{r soc stock methods summary}
soc_method_summary <- esm %>%
  group_by(depth_std, method_longest) %>%
  dplyr::summarize(mean = round(mean(soc_cum),1))

soc_method_minmaxdiff <- esm %>%
  filter(depth_std=="60") %>%
  group_by(project, label, dsp_pedon_id, depth_std) %>%
  dplyr::summarize(min_soc_cum = min(soc_cum),
            max_soc_cum = max(soc_cum),
            diff_soc_cum = max_soc_cum - min_soc_cum,
            pct_diff_soc_cum = ((max_soc_cum - min_soc_cum) / min_soc_cum)*100)

mean_diff_project <- soc_method_minmaxdiff %>%
  ungroup() %>%
  group_by(project) %>%
  dplyr::summarize(mean_diff_soc_cum = round(mean(diff_soc_cum), 2),
            min_diff_soc_cum = round(min(diff_soc_cum), 2),
            max_diff_soc_cum = round(max(diff_soc_cum), 2))

flextable(mean_diff_project)

mean_diff <- soc_method_minmaxdiff %>%
  ungroup() %>%
  dplyr::summarize(mean_diff_soc_cum = round(mean(diff_soc_cum), 2),
            min_diff_soc_cum = round(min(diff_soc_cum), 2),
            max_diff_soc_cum = round(max(diff_soc_cum), 2),
            max_pct_diff_soc_cum = round(max(pct_diff_soc_cum), 1),
            mean_pct_diff_soc_cum = round(mean(pct_diff_soc_cum), 1))

flextable(mean_diff)
```

Try with one-way ANOVA:

```{r aov calculation method}
aov_soc_cum <- esm %>%
  group_by(depth_std) %>%
  nest() %>%
  mutate(aov = purrr::map(data, ~aov(soc_cum ~ method_longest, data = .x)),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(.x = aov, .y = hsd, ~multcompLetters4(.x, .y, reversed=TRUE)),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[1]]$Letters) |> 
             rownames_to_column() |> 
             rename(method_longest = rowname,
                    letter = 2)
         })
  ) |> 
  select(depth_std, letters_df) |> 
  unnest(cols=letters_df)
```

These plots show that (unsurprisingly), using different calculation methods results in significantly different SOC stocks, especially in surface soil horizons. Using ESM min methods results in the smallest stocks (of course), while fixed depth tends to be more comparable to ESM mean estimates.

```{r fig s2 plot actual cumulative SOC stocks at each depth}
fig2_letters_position <- esm %>%
  group_by(depth_std, method_longest) %>%
  dplyr::summarize(max_soc_cum = max(soc_cum)) %>%
  left_join(aov_soc_cum, by=c("depth_std", "method_longest"))

ggplot(esm, aes(x=method_longest, y=soc_cum, fill=method_longest)) +
  geom_point(aes(color=method_longest), position = "jitter", alpha=0.8) +
  geom_boxplot(fatten=1.6, lwd=0.25, outlier.shape=NA) +
  geom_text(data=fig2_letters_position, aes(y=max_soc_cum + 30, x=method_longest, label=letter), size=2) +
  labs(x="Calculation method", y=expression("Cumulative SOC stock (Mg ha"^-1*")")) +
  facet_wrap(~depth_std, labeller=labeller(depth_std=cum_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_color_paletteer_d("nationalparkcolors::Arches",
                          name="Calculation method") + 
  scale_fill_paletteer_d("nationalparkcolors::Arches",
                          name="Calculation method") + 
  theme_katy() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")

# ggsave(here("figs", "revision2_figs", "figsupp2_soc_cum_all.png"), width=120, height=100, units="mm", dpi=400)
```

Also plot individually for each project:

```{r cumulative soc stocks per project}
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=method_longest, y=soc_cum, fill=method_longest)) +
               geom_point(aes(color=method_longest), position = "jitter", alpha=0.8) +
               geom_boxplot(fatten=1.6, lwd=0.25, outlier.shape=NA) +
               labs(x="Calculation method", y="SOC stock (Mg/ha)",
                    title=glue::glue("SOC stocks by calculation method - ", project_labels_esm[.x])) +
               facet_wrap(~depth_std, labeller=labeller(depth_std=cum_depth_labels)) +
               scale_x_discrete(labels=method_labels) +
               scale_color_paletteer_d("nationalparkcolors::Arches",
                                       name="Calculation method") + 
               scale_fill_paletteer_d("nationalparkcolors::Arches",
                                      name="Calculation method") + 
               theme_classic() +
               theme(axis.text.x=element_text(angle=45, hjust=1),
                     legend.position="none")
             
             # ggsave(here("figs", "revision2_figs", "proj_soc_stock", glue::glue("suppfig_stock_", .x, ".png")),
             #                                  width=120, height=100, units="mm", dpi=400)
           })
```

# How does calculation method influence management sensitivity?

## Calculate means for all treatments, depths, and calculation methods for results text

```{r mean SOC stock in all treatments depths and calc methods}
esm |> 
  group_by(method_longest, esm_depth, label) |> 
  dplyr::summarize(mean = round(mean(soc_cum), 1)) |> 
  arrange(esm_depth, label, method_longest) |> 
  flextable()

soc_cum_summary |> 
  group_by(esm_depth, label) |> 
  dplyr::summarize(min = min(mean),
            max = max(mean)) |> 
  flextable()
```

## Management sensitivity of cumulative SOC stocks

ANOVA for management sensitivity of cumulative SOC stocks using different calculation methods

```{r aov management}
# anova on effect of management on cumulative SOC stocks
mgmt_aov_df <- esm |> 
  group_by(method_longest, depth_std) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(soc_cum ~ label, data=.x)),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, ~multcompLetters4(.x, .y, reversed=TRUE)),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[1]]$Letters) |> 
             rownames_to_column() |> 
             rename(label = rowname,
                    letter = 2)
         }) 
  ) |> 
  select(depth_std, method_longest, letters_df) |> 
  unnest(cols=letters_df)

mgmt_aov_letters <- esm %>%
  group_by(method_longest, depth_std) %>%
  dplyr::summarize(max_soc_cum = max(soc_cum)) %>%
  left_join(mgmt_aov_df, by=c("depth_std", "method_longest")) %>%
  mutate(label = factor(label, levels=c("BAU", "SHM", "Ref")))
```

```{r fig 3 soc mgmt aov letters}
ggplot(esm, aes(x=method_longest, y=soc_cum, fill=label)) +
  geom_point(aes(color=label), pch = 21, position=position_jitterdodge(), alpha=0.8, size=0.7) +
  geom_boxplot(fatten=1.6, lwd=0.25, outlier.shape=NA) +
  scale_x_discrete(labels=method_labels) +
  facet_wrap(~depth_std, scales="free", labeller=labeller(depth_std=cum_depth_labels)) +
  geom_text(data=mgmt_aov_letters, 
            aes(y=max_soc_cum + 25, label=letter, group=label), 
            color="black", position=position_dodge(width=0.8), size=2) +
  labs(x="Calculation method", y=expression("Cumulative SOC (Mg ha"^-1*")")) +
  scale_fill_manual(values=mgmt_pal,
                    breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  scale_color_manual(values=mgmt_pal,
                     breaks=c("BAU", "SHM", "Ref"), 
                     name="Management") +
  theme_katy() +
  theme(axis.text.x=element_text(angle=45, hjust=1))
# ggsave(here("figs", "revision2_figs", "fig3_soc_mgmt.png"), width=120, height=100, units="mm", dpi=400)
```

When accounting for differences in depth and project, all calculation methods result in SOC stocks that are highest in Ref condition, and lower in both SHM and BAU conditions.

# Calculation method influence on SOC sequestration

Calculate the mean SOC in the BAU treatment for each project, and then calculate the difference between each Ref/SHM pedon and the BAU mean.

This isn't entirely different from assessing the management sensitivity of each calculation method, as we did above. However, in this analysis we are concerned with the magnitude of the difference in SOC stocks between both Ref and SHM treatments compared to BAU.

```{r calculate difference between BAU and Ref/SHM SOC stocks}
esm_bau <- esm %>%
  filter(label=="BAU") %>%
  group_by(project, method_long, method_longest, method, ref_stat, ref_data, depth_std) %>%
  dplyr::summarize(mean_bau_soil_mass = mean(soil_mass),
            mean_bau_soc = mean(soc),
            mean_bau_soil_mass_cum = mean(soil_mass_cum),
            mean_bau_soc_cum = mean(soc_cum))

esm_diff <- esm %>%
  filter(!label=="BAU") %>%
  left_join(esm_bau, by=c("project", "method_long", "method_longest", "method", "ref_stat", "ref_data", "depth_std")) %>%
  mutate(diff_soil_mass = soil_mass - mean_bau_soil_mass,
         diff_soc = soc - mean_bau_soc,
         diff_soil_mass_cum = soil_mass_cum - mean_bau_soil_mass_cum,
         diff_soc_cum = soc_cum - mean_bau_soc_cum)

# Calculate mean sequestration estimates for each treatment, depth, and method
esm_diff |> 
  group_by(label, depth_std, method_longest) |> 
  dplyr::summarize(mean_diff = round(mean(diff_soc_cum), 1),
            n = n()) |> 
  flextable()
```

## Cumulative SOC sequestration with different calculation methods

```{r aov calc method dsoc}
dsoc_aov_df <- esm_diff |> 
  group_by(label, depth_std) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(diff_soc_cum ~ method_longest, data=.x)),
         tidy_aov = purrr::map(aov, broom::tidy),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, ~multcompLetters4(.x, .y, reversed=TRUE)),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[1]]$Letters) |> 
             rownames_to_column() |> 
             rename(method_longest = rowname,
                    letter = 2)
         }) 
  )

dsoc_aov_tidy <- dsoc_aov_df |> 
  select(depth_std, label, tidy_aov) |> 
  unnest(cols=tidy_aov) |> 
  mutate(comparison = "dsoc") |> 
  relocate(comparison)

dsoc_letters_df <- dsoc_aov_df |> 
  select(depth_std, label, letters_df) |> 
  unnest(cols=letters_df)
# no significant differences
```

```{r fig4 dsoc}
ggplot(esm_diff, aes(x=method_longest, y=diff_soc_cum, fill=method_longest)) +
  geom_point(aes(color=method_longest), pch=21, position="jitter", alpha=0.8, size=0.8) +
  geom_boxplot(fatten=1.6, lwd=0.25, outlier.shape=NA) +
  geom_abline(intercept=0, slope=0, linetype="dashed", linewidth=0.3) +
  labs(x="Calculation method", y=expression(Delta*"SOC vs BAU (Mg ha"^-1*")")) +
  facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=cum_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_fill_paletteer_d("nationalparkcolors::Arches",
                         name="Calculation method") + 
  scale_color_paletteer_d("nationalparkcolors::Arches",
                         name="Calculation method") + 
  theme_katy() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")
# ggsave(here("figs", "revision2_figs", "fig4_dsoc.png"), width=140, height=120, units="mm", dpi=400)
```

Also plot for each project:

```{r dsoc for each project}
purrr::map(.x = projects, 
           .f = ~{
             esm_diff %>%
               filter(project==.x) %>%
               ggplot(aes(x=method_longest, y=diff_soc_cum, fill=method_longest)) +
               geom_point(aes(color=method_longest), pch=21, position="jitter", alpha=0.8, size=0.8) +
               geom_boxplot(fatten=1.6, lwd=0.25, outlier.shape=NA) +
               geom_abline(intercept=0, slope=0, linetype="dashed", linewidth=0.3) +
               labs(x="Calculation method", y=expression(Delta*"SOC vs BAU (Mg/ha)"),
                    title=glue::glue("\u0394SOC by calculation method - {project_labels_esm[.x]}")) +
               facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=cum_depth_labels)) +
               scale_x_discrete(labels=method_labels) +
               scale_fill_paletteer_d("nationalparkcolors::Arches",
                                      name="Calculation method") + 
               scale_color_paletteer_d("nationalparkcolors::Arches",
                                       name="Calculation method") + 
               theme_classic() +
               theme(axis.text.x=element_text(angle=45, hjust=1),
                     legend.position="none")
             
             # ggsave(here("figs", "revision2_figs", "proj_dsoc", glue::glue("suppfig_dsoc_", .x, ".png")),
             #                                  width=140, height=120, units="mm", dpi=400)
           })
```

# Can we calculate the error in the delta SOC associated with different methods?

```{r delta soc error}
# think we have to assume that something is the true value
# let's say ESMproject,min - whatever is the lowest reference value is the true reference value
# for each point, the error is the difference between delta SOC in that method and delta SOC with ESMproject,min

true_delta_soc <- esm_diff |> 
  filter(method_longest=="esm2_min_project") |> 
  distinct() |> 
  dplyr::select(project, dsp_pedon_id, layer, depth_std, diff_soc_cum) |> 
  rename(true_delta_soc = diff_soc_cum)

esm_error <- esm_diff |> 
  left_join(true_delta_soc, by=c("project", "dsp_pedon_id", "layer", "depth_std")) |> 
  mutate(delta_soc_error = diff_soc_cum - true_delta_soc,
         abs_error = abs(delta_soc_error),
         pct_error = abs((diff_soc_cum - true_delta_soc)/true_delta_soc)) |> 
  filter(method_longest!="esm2_min_project")
```

Plot dSOC percent error for each method:

```{r aov delta soc error}
# anova on effect of method on percent error
dsoc_error_aov_df <- esm_error |> 
  group_by(depth_std, label) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(pct_error ~ method_longest, data=.x)),
         tidy_aov = purrr::map(aov, broom::tidy),
         glance_aov = purrr::map(aov, broom::glance),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, multcompLetters4),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[1]]$Letters) |> 
             rownames_to_column() |> 
             rename(method_longest = rowname,
                    letter = 2)
         }) 
  )

dsoc_error_aov_tidy <- dsoc_error_aov_df |> 
  select(depth_std, label, tidy_aov) |> 
  unnest(cols=tidy_aov) |> 
  mutate(comparison = "dsoc_error") |> 
  relocate(comparison)

dsoc_error_letters_df <- dsoc_error_aov_df |> 
  select(depth_std, label, letters_df) |> 
  unnest(cols=letters_df)

# calculate position
dsoc_error_quant <- esm_error |> 
  group_by(depth_std, method_longest, label) |> 
  summarize(quant25 = quantile(pct_error, probs = 0.25), 
            quant75 = quantile(pct_error, probs = 0.75)) |> 
  mutate(iqr = quant75-quant25,
         upr = quant75 + (1.5*iqr))

dsoc_error_letters_df2 <- dsoc_error_letters_df |> 
  left_join(dsoc_error_quant, by=c("method_longest", "depth_std", "label"))

esm_error_sub <- esm_error |> 
  left_join(select(dsoc_error_quant, upr, depth_std, method_longest), by=c("depth_std", "method_longest")) |> 
  filter(pct_error<=upr)
```

```{r fig 5 dsoc error}
ggplot(esm_error_sub, aes(x=method_longest, y=pct_error, fill=method_longest)) +
  geom_point(aes(color=method_longest), pch=21, position="jitter", alpha=0.8, size=0.8) +
  geom_boxplot(fatten=1.6, lwd=0.25, outliers=FALSE) +
  geom_text(data=dsoc_error_letters_df2, aes(y=upr, label=letter), vjust=-0.6, size=8/.pt) +
  labs(x="Calculation method", y=expression(Delta*"SOC error (%)")) +
  facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=cum_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = c(.05, .2))) +
  scale_fill_paletteer_d("nationalparkcolors::Arches",
                         name="Calculation method") +
  scale_color_paletteer_d("nationalparkcolors::Arches",
                          name="Calculation method") +
  theme_katy() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")
# ggsave(here("figs", "revision2_figs", "fig5_dsoc_error.png"), width=140, height=120, units="mm", dpi=400)
```

Make ANOVA table for text:

```{r dsoc and dsoc error anova table}
anova_table <- bind_rows(dsoc_aov_tidy, dsoc_error_aov_tidy) |> 
  select(-sumsq, -meansq) |> 
  mutate(statistic = round(statistic, 2),
         p.value = if_else(p.value < 0.001, "<0.001", as.character(round(p.value, 3)))) 
anova_table

# write_csv(anova_table, here("figs", "revision2_figs", "dsoc_anova_table.csv"))
```

Absolute value of error on meta-project level:

```{r absolute dsoc error anova}
# anova on effect of method on percent error
dsoc_abserror_aov_df <- esm_error |> 
  group_by(depth_std, label) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(abs_error ~ method_longest, data=.x)),
         tidy_aov = purrr::map(aov, broom::tidy),
         glance_aov = purrr::map(aov, broom::glance),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, multcompLetters4),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[1]]$Letters) |> 
             rownames_to_column() |> 
             rename(method_longest = rowname,
                    letter = 2)
         }) 
  )

dsoc_abserror_aov_tidy <- dsoc_abserror_aov_df |> 
  select(depth_std, label, tidy_aov) |> 
  unnest(cols=tidy_aov) |> 
  mutate(comparison = "dsoc_abserror") |> 
  relocate(comparison)

dsoc_abserror_letters_df <- dsoc_abserror_aov_df |> 
  select(depth_std, label, letters_df) |> 
  unnest(cols=letters_df)

# calculate position
dsoc_abserror_quant <- esm_error |> 
  group_by(depth_std, method_longest, label) |> 
  summarize(quant25 = quantile(abs_error, probs = 0.25), 
            quant75 = quantile(abs_error, probs = 0.75)) |> 
  mutate(iqr = quant75-quant25,
         upr = quant75 + (1.5*iqr))

dsoc_abserror_letters_df2 <- dsoc_abserror_letters_df |> 
  left_join(dsoc_abserror_quant, by=c("method_longest", "depth_std", "label"))
```

```{r absolute dsoc error plot}
# filter to just points within 1.5*IQR for ease of plotting
esm_abserror_sub <- esm_error |> 
  left_join(select(dsoc_abserror_quant, upr, depth_std, method_longest), by=c("depth_std", "method_longest")) |> 
  filter(abs_error<=upr)

ggplot(esm_error_sub, aes(x=method_longest, y=abs_error, fill=method_longest)) +
  geom_point(aes(color=method_longest), pch=21, position="jitter", alpha=0.8, size=0.8) +
  geom_boxplot(fatten=1.6, lwd=0.25, outliers=FALSE) +
  geom_text(data=dsoc_abserror_letters_df2, aes(y=upr, label=letter), vjust=-0.8, size=8/.pt) +
  labs(x="Calculation method", y=expression(Delta*"SOC -"~Delta*"SOC ESM"[("project,"~min)]~"(Mg ha"^-1*")")) +
  facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=cum_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_y_continuous(expand = expansion(mult = c(.05, .2))) +
  scale_fill_paletteer_d("nationalparkcolors::Arches",
                         name="Calculation method") +
  scale_color_paletteer_d("nationalparkcolors::Arches",
                          name="Calculation method") +
  theme_katy() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")
# ggsave(here("figs", "revision2_figs", "fig5_dsoc_abserror.png"), width=140, height=120, units="mm", dpi=400)
```

dSOC error for each project for supplement:

```{r dsoc error for each project}
purrr::map(.x = projects, 
           .f = ~{
             esm_error %>%
               filter(project==.x) %>%
               ggplot(aes(x=method_longest, y=delta_soc_error, fill=method_longest)) +
               geom_boxplot(fatten=1.6, lwd=0.25, outliers=FALSE) +
               geom_abline(intercept=0, slope=0, linetype="dashed", linewidth=0.3) +
                labs(x="Calculation method", y=expression(Delta*"SOC error (Mg ha"^-1*")"),
                    title=glue::glue("\u0394SOC error by calculation method - {project_labels_esm[.x]}")) +
               facet_wrap(~depth_std, labeller=labeller(depth_std=cum_depth_labels)) +
               scale_x_discrete(labels=method_labels) +
               scale_fill_paletteer_d("nationalparkcolors::Arches",
                                      name="Calculation method") + 
               theme_katy() +
               theme(axis.text.x=element_text(angle=45, hjust=1),
                     legend.position="none")
             
             # ggsave(here("figs", "revision2_figs", "proj_dsoc_error", glue::glue("suppfig_dsoc_error", .x, ".png")),
             #                                  width=120, height=100, units="mm", dpi=400)
           })
```

Table of means for text:

```{r dsoc error table}
esm_error |> 
  group_by(depth_std, method_longest) |> 
  summarize(mean_pct_error = round(mean(pct_error), 3) * 100) |> 
  flextable()
```

## What drives dSOC error?

### Relationship between bulk density variability and dSOC error

Try calculating the percent difference between an individual pedon's BD in each depth increment and the mean BAU depth increment BD for that project:

```{r calculate pct diff bd}
# first need to calculate pct difference in bulk density between treatments for each depth increment
bd_mean_bau <- bd_depth |> 
  select(-depth) |> 
  rename(depth_std = layer) |> 
  group_by(project, label, depth_std) |> 
  summarize(bd_mean_bau = mean(depth_wt_bd)) |> 
  filter(label=="BAU") |> 
  ungroup() |> 
  select(-label)

bd_error <- bd_depth |> 
  select(-depth) |> 
  rename(depth_std = layer) |> 
  filter(label!="BAU") |> 
  left_join(bd_mean_bau, by=c("project", "depth_std")) |> 
  mutate(bd_diff = depth_wt_bd - bd_mean_bau,
         abs_bd_error = abs(depth_wt_bd - bd_mean_bau),
         pct_error_bd = abs_bd_error / bd_mean_bau)

# then need to calculate dSOC error in each depth increment
true_inc_delta_soc <- esm_diff |> 
  filter(method_longest=="esm2_min_project") |> 
  distinct() |> 
  dplyr::select(project, dsp_pedon_id, layer, depth_std, diff_soc) |> 
  rename(true_inc_delta_soc = diff_soc)

esm_inc_error <- esm_diff |> 
  left_join(true_inc_delta_soc, by=c("project", "dsp_pedon_id", "layer", "depth_std")) |> 
  mutate(inc_delta_soc_error = diff_soc - true_inc_delta_soc,
         abs_error = abs(inc_delta_soc_error),
         pct_error = abs((diff_soc - true_inc_delta_soc)/true_inc_delta_soc)) |> 
  filter(method_longest!="esm2_min_project")
```

```{r dsoc error vs bd error regression table}
esm_inc_error_bd_error <- esm_inc_error |> 
  left_join(select(bd_error, dsp_pedon_id, depth_std, bd_diff, abs_bd_error, pct_error_bd), by=c("dsp_pedon_id", "depth_std")) |> 
  mutate(abs_true_inc_delta_soc = abs(true_inc_delta_soc))

# all regressions
dsoc_bd_error_lm <- esm_inc_error_bd_error |> 
  group_by(method_longest, depth_std) |> 
  nest() |> 
  mutate(lm = purrr::map(data, ~lm(inc_delta_soc_error ~ bd_diff, data=.x)),
         lm_tidy = purrr::map(lm, broom::glance))

dsoc_bd_error_lm_table <- dsoc_bd_error_lm |> 
  select(method_longest, depth_std, lm_tidy) |> 
  unnest(cols=lm_tidy) |> 
  select(method_longest, depth_std, statistic, df, df.residual, p.value, r.squared) |> 
  mutate(statistic = round(statistic, 2),
         r.squared = round(r.squared, 2),
         p.value = ifelse(p.value < 0.001, "<0.001", as.character(round(p.value, 3))))
dsoc_bd_error_lm_table
# write_csv(dsoc_bd_error_lm_table, here("figs", "revision2_figs", "dsoc_bd_regression_table.csv"))
```

```{r sig error dsoc and bd regressions in fd}
fd_error <- esm_inc_error_bd_error |> 
  filter(method_longest == "fd_fd_fd")

# first figure out which regressions are significant
fd_reg <- fd_error %>%
  group_by(method_longest, depth_std) %>%
  nest() %>%
  mutate(lm = purrr::map(data, ~lm(inc_delta_soc_error ~ bd_diff, data=.x)),
         lm_tidy = purrr::map(lm, tidy),
         lm_adjr2 = purrr::map(lm, .f = ~{
           glance(.x) %>%
             select(adj.r.squared)
           }))

fd_sig_reg <- fd_reg %>%
  unnest(cols=c(lm_tidy, lm_adjr2)) %>%
  filter(term=="bd_diff") %>%
  filter(p.value < 0.05) %>%
  unite("method_depth", c("method_longest", "depth_std"), remove=FALSE) %>%
  mutate(p_print = if_else(p.value < 0.001, "<0.001", paste0("=='", format(round(p.value, 3), nsmall = 3), "'"))) %>%
  mutate(reg_lab = paste0("~R^{2} == ", round(adj.r.squared, 2), "*`,`~p", p_print))
```

```{r fig 6 dsoc error vs bd error fixed depth}
ggplot(fd_error, aes(x=bd_diff, y=inc_delta_soc_error)) +
  geom_point(aes(color=project, shape=label)) +
  geom_abline(slope=0, intercept=1, linetype="dashed") +
  geom_vline(xintercept=0, linetype="dashed") +
  geom_smooth(method="lm", se=FALSE, color="black") +
  geom_text(data=fd_sig_reg,
           aes(label = reg_lab, y=50, x=-0.7),
           size = 7.5/.pt, parse=TRUE) +
  facet_wrap(~depth_std, labeller=labeller(depth_std = inc_depth_labels)) +
  labs(x =expression("Management bulk density difference (g cm"^-3*")"), 
       y = expression("Fixed depth"~Delta*"SOC error (Mg ha"^-1*")")) +
  scale_color_manual(values=pnw_palette("Sunset2", 9), name="Project",
                     labels = project_labels_esm) + 
  scale_shape_discrete(name="Management") + 
  theme_katy()
# ggsave(here("figs", "revision2_figs", "fig6_fd_dsoc_error_vs_bd_error.png"), width=150, height=110, units="mm", dpi=400)
```

Also show dSOC error vs BD error for ESM methods:

```{r fig s12 dsoc error vs bd error for esm methods}
# other methods - fig for supplement

# first figure out which regressions are significant
regressions <- esm_inc_error_bd_error %>%
  filter(method_longest!="fd_fd_fd") %>%
  group_by(method_longest, depth_std) %>%
  nest() %>%
  mutate(lm = purrr::map(data, ~lm(inc_delta_soc_error ~ bd_diff, data=.x)),
         lm_tidy = purrr::map(lm, tidy),
         lm_adjr2 = purrr::map(lm, .f = ~{
           glance(.x) %>%
             select(adj.r.squared)
           }))

sig_reg <- regressions %>%
  unnest(cols=c(lm_tidy, lm_adjr2)) %>%
  filter(term=="bd_diff") %>%
  filter(p.value < 0.05) %>%
  unite("method_depth", c("method_longest", "depth_std"), remove=FALSE) %>%
  mutate(p_print = if_else(p.value < 0.001, "<0.001", paste0("=='", format(round(p.value, 3), nsmall = 3), "'"))) %>%
  mutate(reg_lab = paste0("~R^{2} == ", round(adj.r.squared, 2), "*`,`~p", p_print))

error_methods_to_plot <- esm_inc_error_bd_error |> 
  filter(method_longest!="fd_fd_fd") |> 
  unite("method_depth", c("method_longest", "depth_std"), remove=FALSE)

ggplot(error_methods_to_plot, aes(x=bd_diff, y=inc_delta_soc_error)) +
  geom_point(aes(color=project)) +
  geom_abline(slope=0, intercept=1, linetype="dashed") +
  geom_vline(xintercept=0, linetype="dashed") +
  geom_smooth(data=error_methods_to_plot |>  filter(method_depth %in% sig_reg$method_depth),
                           method="lm", color="black", lwd=0.5, se=FALSE) +
  geom_text(data=sig_reg,
           aes(label = reg_lab, y=50, x=-0.4),
           size = 7.5/.pt, parse=TRUE) +
  facet_grid(method_longest~depth_std,
             labeller=labeller(depth_std = inc_depth_labels,
                               method_longest = method_labels2)) +
  labs(x =expression("Management bulk density difference (g cm"^-3*")"), 
       y = expression(Delta*"SOC error (Mg ha"^-1*")")) +
  scale_color_manual(values=pnw_palette("Sunset2", 9), name="Project",
                     labels = project_labels_esm) +  
  theme_katy()
# ggsave(here("figs", "revision2_figs", "figsupp3_all_methods_dsoc_error_vs_bd_error.png"), width=180, height=140, units="mm", dpi=400)
```

Error in FD dSOC compared to ESM(min) dSOC is significantly correlated with the difference between sample BD and mean BAU bd - in other words, when bulk density changes between treatments, 

### Relationship between true dSOC and dSOC error

```{r fig s13 dsoc error vs dsoc}
# first figure out which regressions are significant
dsoc_regressions <- esm_inc_error_bd_error |> 
  group_by(method_longest, depth_std) |> 
  nest() |> 
  mutate(lm = purrr::map(data, ~lm(inc_delta_soc_error ~ true_inc_delta_soc, data=.x)),
         lm_tidy = purrr::map(lm, tidy),
         lm_adjr2 = purrr::map(lm, .f = ~{
           glance(.x) |> 
             select(adj.r.squared)
           }))

dsoc_sig_reg <- dsoc_regressions |> 
  unnest(cols=c(lm_tidy, lm_adjr2)) |> 
  filter(term=="true_inc_delta_soc") |> 
  filter(p.value < 0.05) |> 
  unite("method_depth", c("method_longest", "depth_std"), remove=FALSE) |> 
  mutate(p_print = if_else(p.value < 0.001, "<0.001", paste0("=='", format(round(p.value, 3), nsmall = 3), "'"))) |> 
  mutate(reg_lab = paste0("~R^{2} == ", round(adj.r.squared, 2), "*`,`~p", p_print))

dsoc_to_plot <- esm_inc_error_bd_error |> 
  unite("method_depth", c("method_longest", "depth_std"), remove=FALSE)

# plot error against true delta SOC
ggplot(dsoc_to_plot, aes(x=true_inc_delta_soc, y=inc_delta_soc_error)) +
  geom_point(aes(color=project)) +
  geom_hline(yintercept=0, linetype="dashed") +
  geom_vline(xintercept=0, linetype="dashed") +
  geom_smooth(data=dsoc_to_plot |>  filter(method_depth %in% sig_reg$method_depth),
                           method="lm", color="black", lwd=0.5, se=FALSE) +
  geom_text(data=dsoc_sig_reg,
           aes(label = reg_lab, y=50, x=10),
           size = 7/.pt, parse=TRUE) +
  labs(x = expression(Delta*"SOC ESM"["(project, min)"]~"(Mg ha"^-1*")"),
       y = expression(Delta*"SOC error (Mg ha"^-1*")")) +
  facet_grid(method_longest~depth_std, labeller = labeller(depth_std = inc_depth_labels,
                               method_longest = method_labels2)) +
  scale_color_manual(values=pnw_palette("Sunset2", 9), name="Project",
                     labels = project_labels_esm) +  
  theme_katy()
# ggsave(here("figs", "revision2_figs", "figsupp4_dsocerror_vs_truedsoc.png"), width=180, height=140, units="mm", dpi=400)
```

# Check that ESM calculations worked and make sense

## Plot mean cumulative SOC vs depth, and cumulative SOC vs cumulative soil mass

These plots are just to check that the ESM calculations worked and nothing looks too crazy - for example, wild fluctuations in SOC content with depth, impossible numbers, etc.

Cumulative mass vs depth:

```{r cumulative mass vs depth}
ggplot(esm, aes(x=depth_std, y=soil_mass_cum, linetype=method_longest, color=project)) +
  stat_summary(geom="line", fun="mean", linewidth=1.3) +
  coord_flip() +
  xlim(c(60,0)) +
  labs(x="ESM Depth (cm)", y="Cumulative SOC (Mg/ha)", 
       title="Cumulative mass with depth") +
  scale_color_manual(values=pnw_palette("Sunset2", 9), name="Project",
                     labels = project_labels_esm) +  
  theme_classic()
```

Cumulative SOC vs depth:

```{r cumulative soc vs depth}
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=depth_std, y=soc_cum, color=method_longest)) +
               stat_summary(geom="line", fun="mean", linewidth=1.3) +
               coord_flip() +
               xlim(c(60,0)) +
               facet_grid(~label, scales="free") +
               labs(x="ESM Depth (cm)", y="Cumulative SOC (Mg/ha)", 
                    title=glue::glue("Cumulative SOC with depth - ", .x)) +
               scale_color_paletteer_d("nationalparkcolors::Arches",
                                       name="Calculation method", labels=method_labels) + 
               theme_classic()
           })
```
 
These aren't the easiest to interpret, but I think that everything looks good here. It doesn't look like SOC estimates take a crazy left turn at any depth for any of the methods, and they are all in relative agreement with each other. The plot of cumulative mass vs cumulative SOC shows that using the minimum reference mass for ESM results in a lower cumulative soil mass overall, but that's no surprise. Fixed depth results look pretty identical to ESM mean.
 
## How does ESM choice influence soil mass in each depth increment?

This is another way to check that nothing is getting wonky with the ESM calculations - making sure that the soil mass in each depth increment makes sense for each project.
 
 Plot soil mass in each depth increment:
 
```{r soil mass in each depth increment} 
# Soil mass in each depth increment
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=esm_depth, y=soil_mass, fill=method_longest)) +
               stat_summary(geom="bar", fun=mean, position="dodge2", color="black") +
               stat_summary(geom="errorbar", fun.data = mean_cl_normal, color="black",
                            width=0.4, position=position_dodge(width=0.9)) +
               coord_flip() +
               scale_x_discrete(limits=rev) +
               facet_wrap(~label) +
               labs(x="ESM Depth", y="Depth Increment Soil Mass (Mg/ha)", 
                    title=glue::glue("Depth increment soil mass by calculation method - ", .x)) +
               scale_fill_paletteer_d("nationalparkcolors::Arches",
                                      name="Calculation method",
                                      labels=method_labels) + 
               theme_classic()
           })
```

Plot cumulative soil mass in each depth increment:

```{r cumulative soil mass in each depth increment}
# Cumulative soil mass in each depth increment
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=factor(depth_std), y=soil_mass_cum, fill=method_longest)) +
               stat_summary(geom="bar", fun=mean, position="dodge2", color="black") +
               stat_summary(geom="errorbar", fun.data=mean_cl_normal, color="black",
                            width=0.4, position=position_dodge(width = 0.9)) +
               coord_flip() +
               scale_x_discrete(limits=rev) +
               facet_wrap(~label) +
               labs(x="ESM Depth (cm)", y="Cumulative Soil Mass (Mg/ha)", 
                    title=glue::glue("Cumulative soil mass by calculation method - ", .x)) +
               scale_fill_paletteer_d("nationalparkcolors::Arches",
                                      name="Calculation method",
                                      labels=method_labels) + 
               theme_classic()
           })
```

These plots are a check to see how using ESM methods influence the mass of soil in a given depth increment. They show that, unsurprisingly, fixed depth calculations result in the greatest soil mass in each depth increment. Fixed depth soil masses are most comparable to ESM(mean) masses.