---
title: "esm_manuscript_revision2"
author: "Katy Dynarski"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(here)
library(janitor)
library(paletteer)
library(aqp)
library(broom)
library(broom.mixed)
library(lme4)
library(lmerTest)
library(multcomp)
library(ggeffects)
library(ggpubr)
library(sf)
library(ggspatial)
library(ggeasy)
library(ggrepel)
library(patchwork)
library(multcomp)
library(multcompView)
library(PNWColors)
library(DBI)
library(RSQLite)
library(soilDB)
library(tidyverse)

# Data
# Vector for method order for consistent plotting
method_factor_order <- c("esm2_min_project", "esm2_min_treatment", "esm2_mean_project", "esm2_mean_treatment", 
                         "esm2_max_project", "esm2_max_treatment", "fd_fd_fd")

method_factor_order2 <- c("min_project", "min_treatment", "mean_project", "mean_treatment", 
                         "max_project", "max_treatment", "fd_fd")

# Vector for project order for consistent plotting
project_plotting_order <- c("UConn", "Illinois", "KansasState", "UnivOfMinnesota", "NCState", 
                            "OregonStateJory", "OregonStateWoodburn","TexasA&MPt-1", "WashingtonState")

esm_all <- read.csv(here("data_processed", "esm_revision3", "esm_standard_depths.csv")) |>
  filter(project!="UTRGV") |> # don't include UTRGV for this analysis - bulk density is from SSURGO
  unite("method_long", c("method", "ref_stat"), sep="_", remove = FALSE) |>
  unite("method_longest", c("method", "ref_stat", "ref_data"), sep="_", remove = FALSE) |>
  mutate(depth_std = case_when(layer==1 ~ 5,
                               layer==2 ~ 10,
                               layer==3 ~ 30,
                               layer==4 ~ 60),
         esm_depth = factor(esm_depth, levels=c("0-5 cm", "5-10 cm", "10-30 cm", "30-60 cm")),
         label = factor(label, levels=c("BAU", "SHM", "Ref")),
         project = factor(project, levels=project_plotting_order))

esm <- esm_all |> 
  filter(method!="esm1") |> # not including ESM classic in this analysis - just compare cubic spline
  mutate(method_longest = factor(method_longest, levels=method_factor_order))
  
esm1 <- esm_all |> 
  filter(method=="esm1")
  
horizon_data <- read.csv(here("data_processed", "soc_horizon_esm_input.csv"))
project <- read.csv(here("data_processed", "project_data_meta.csv"))

# List of projects
projects <- esm |> 
  distinct(project) |> 
  mutate(project = as.character(project)) |>
  pull()

inc_depth_labels <- c(
  "5" = "ESM Increment\nDepth: 0-5 cm",
  "10" = "ESM Increment\nDepth: 5-10 cm",
  "30" = "ESM Increment\nDepth: 10-30 cm",
  "60" = "ESM Increment\nDepth: 30-60 cm"
)

cum_depth_labels <- c(
  "5" = "ESM Cumulative\nDepth: 0-5 cm",
  "10" = "ESM Cumulative\nDepth: 0-10 cm",
  "30" = "ESM Cumulative\nDepth: 0-30 cm",
  "60" = "ESM Cumulative\nDepth: 0-60 cm"
)
  
method_labels <- c(
  "esm2_min_project" = expression("ESM"["project, min"]),      
  "esm2_mean_project" = expression("ESM"["project, mean"]),
  "esm2_max_project" = expression("ESM"["project, max"]),
  "esm2_min_treatment"= expression("ESM"["control, min"]),   
  "esm2_mean_treatment" = expression("ESM"["control, mean"]),
  "esm2_max_treatment" = expression("ESM"["control, max"]),
  "fd_fd_fd" = "Fixed depth"
)

method_labels2 <- c(
  "esm2_min_project" = "ESM(project, min)",      
  "esm2_mean_project" = "ESM(project, mean)", 
  "esm2_max_project" = "ESM(project, max)",
  "esm2_min_treatment"= "ESM(control, min)",   
  "esm2_mean_treatment" = "ESM(control, mean)",
  "esm2_max_treatment" = "ESM(control, max)",
  "fd_fd_fd" = "Fixed depth"
)

stat_data_labels <- c(
  "min_project" = expression("ESM"["project, min"]),      
  "mean_project" = expression("ESM"["project, mean"]),
  "max_project" = expression("ESM"["project, max"]),
  "min_treatment"= expression("ESM"["control, min"]),   
  "mean_treatment" = expression("ESM"["control, mean"]),
  "max_treatment" = expression("ESM"["control, max"]),
  "fd_fd" = "Fixed depth"
)

stat_data_labels2 <- c(
  "min_project" = "ESM(project, min)",      
  "mean_project" = "ESM(project, mean)", 
  "max_project" = "ESM(project, max)",
  "min_treatment"= "ESM(control, min)",   
  "mean_treatment" = "ESM(control, mean)",
  "max_treatment" = "ESM(control, max)",
  "fd_fd" = "Fixed depth"
)

project_labels_esm <- c("UnivOfMinnesota" = "U. Minnesota",
                    "WashingtonState" = "Washington State",
                    "TexasA&MPt-1" = "Texas A&M",
                    "NCState" = "North Carolina State",
                    "Illinois" = "U. Illinois",
                    "OregonStateJory" = "Oregon State - Jory",
                    "OregonStateWoodburn" = "Oregon State - Woodburn",
                    "KansasState" = "Kansas State",
                    "UConn" = "U. Connecticut")

project_labels_esm_df <- data.frame(project_labels_esm) |>
  rownames_to_column()

mgmt_pal <- c("#FED789FF","#72874EFF","#476F84FF")

method_pal <- c(esm2_min_project = "#e69b99", 
                esm2_min_treatment = "#ba7999", 
                esm2_mean_project = "#89689d", 
                esm2_mean_treatment = "#59629b",
                esm2_max_project = "#2c6184",
                esm2_max_treatment = "#015b58",
                fd_fd_fd = "#24492e")

method_pal2 <- c(min_project = "#e69b99", 
                min_treatment = "#ba7999", 
                mean_project = "#89689d", 
                mean_treatment = "#59629b",
                max_project = "#2c6184",
                max_treatment = "#015b58",
                fd_fd = "#24492e")

theme_katy <- function(base_size=8) {
  theme_classic(base_size=base_size) %+replace%
    theme(# Legend
      legend.title=element_text(size=rel(1)), legend.text=element_text(size=rel(.8)),
      #axes                      
      axis.text=element_text(size=rel(.8)),
      axis.title=element_text(size=rel(1)),
      axis.line=element_line(linewidth=0.25),
      axis.ticks=element_line(linewidth=0.25),
      strip.background=element_rect(linewidth=0.3),)
}
```

# Overview

## Calculation method comparison

I calculated SOC stocks in 0-5 cm, 5-10 cm, 10-30 cm, and 30-60 cm depth increments via fixed depth and ESM methods. For the ESM calculations, I tested four different reference mass options:

* Minimum soil mass in an individual DSP4SH project (ESM project, min)
* Mean soil mass in an individual DSP4SH project (ESM project, mean)
* Minimum soil mass in the control group (Ref) in an individual DSP4SH project (ESM control, min)
* Mean soil mass in the control group (Ref) in an individual DSP4SH project (ESM control, mean)

# Climate and soils data for each project 

```{r climate means for each project}
# need to rename OSU projects in "project" dataframe and calculate mean MAT/MAP for all projects
project_renamed <- project |>
  mutate(project = case_when(
    project=="OregonState" & soil=="Jory" ~ "OregonStateJory",
    project=="OregonState" & soil=="Woodburn" ~ "OregonStateWoodburn", 
    .default = project))

project_avg_climate <- project_renamed |>
  group_by(project) |>
  dplyr::summarize(avg_mat = round(mean(mat, na.rm=TRUE), 1),
         avg_ppt = round(mean(map, na.rm=TRUE), 0)) |>
  arrange(factor(project, levels=project_plotting_order))

# write_csv(project_avg_climate, here("figs", "revision_figs", "tablesupp1_site_clim.csv"))
```

## Make map
```{r map}
# calculate average lat and long for each project
proj_loc <- project_renamed |> 
  filter(project!="UTRGV", project !="TexasA&MPt-2") |> 
  group_by(project) |> 
  summarize(mean_x = mean(pedon_x), 
            mean_y = mean(pedon_y)) |> 
  left_join(project_labels_esm_df, by=c("project" = "rowname"))

# Download map of USA from maps() and convert to sf
usa <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))

# Make map
ggplot(data=usa) +
  geom_sf(fill=NA) +
  coord_sf(xlim=c(-125.0, -66.93457), ylim=c(23.5, 49.384358)) + # set bounding box around CONUS
  annotation_north_arrow(location="bl", which_north="true", height=unit(.25, "in"), width=unit(.25, "in"),
                         pad_x = unit(0.4, "in"), pad_y = unit(0.25, "in"), style=north_arrow_fancy_orienteering) + # add north arrow
  annotation_scale(location = "bl") +
  geom_point(data=proj_loc, aes(x=mean_x, y=mean_y, fill=project), pch = 21, size = 4) +
  geom_label_repel(data=proj_loc, aes(x=mean_x, y=mean_y, label=project_labels_esm),
                   min.segment.length = 0, seed = 42, box.padding = 0.5) +
  scale_fill_manual(values=pnw_palette("Sunset2", 9)) +
  theme_katy() +
  easy_remove_axes() +
  theme(legend.position="none")
# ggsave(here("figs", "revision3_figs", "figsupp1_map.png"), width=180, height=130, units = "mm")
```

# Number of pedons and sites

```{r how many pedons and sites}
# how many projects
n_distinct(esm$project)

# how many pedons
n_distinct(esm$dsp_pedon_id)

# how many sites
# have to use some regex to extract the site name from the pedon ID since I'm missing the needed column
sites <- esm |>
  dplyr::select(project, label, dsp_pedon_id) |>
  distinct(project, label, dsp_pedon_id) |>
  mutate(dsp_site_id = str_remove(dsp_pedon_id, "-\\d*\\w*$")) |>
  group_by(project, label, dsp_site_id)

# how many pedons per site?
sites |>
  dplyr::summarize(n_pedons_per_site = n())

# how many total sites  
n_distinct(sites$dsp_site_id)

# how many total pedons
n_distinct(sites$dsp_pedon_id)

# how many total pedons per treatment
esm |>
  dplyr::select(project, label, dsp_pedon_id) |>
  distinct(project, label, dsp_pedon_id) |>
  group_by(label) |>
  dplyr::summarize(n_pedons_per_treatment = n())

# table for supplement with number of sites and pedons per treatment in each project
site_table <- sites |>
  distinct(project, label, dsp_site_id) |>
  group_by(project, label) |>
  dplyr::summarize(n_sites_per_treatment = n())

pedon_table <- sites |>
  ungroup() |>
  group_by(project, label) |>
  dplyr::summarize(n_pedons_per_trt = n())

site_pedon_table <- site_table |>
  left_join(pedon_table, by=c("project", "label")) |>
  arrange(factor(project, levels=project_plotting_order), label)

# write_csv(site_pedon_table, here("figs", "revision_figs", "tablesupp2_site_pedon_numbers.csv"))
```

# Range in bulk density and reference masses

Management can influence soil bulk density, which results in different masses of soil within the same depth increment between treatments. We can check this by plotting soil bulk density under the different management treatments in each project.

Plot bulk density:

```{r calculate bulk density by depth increment, include=FALSE}
bd <- horizon_data |>
  filter(!project=="UTRGV") |>
  dplyr::select(project, label, dsp_pedon_id, dsp_sample_id, hrzdep_t, hrzdep_b, bd_hybrid)

# promote to SPC and dice into 1-cm increments
spc <- bd
aqp::depths(spc) <- dsp_pedon_id ~ hrzdep_t + hrzdep_b

dice <- aqp::dice(spc, fm=0:99 ~ bd_hybrid + project + label)
dice_bd <- horizons(dice)

bd_depth <- c(5, 10, 30, 60, 100)

# Initialize output vector
out <- vector("list", length(bd_depth))

for (i in seq_along(bd_depth)) {
    if (i == 1) {
      out[[i]] <- dice_bd |>
        filter(hrzdep_t < bd_depth[[i]]) |>
        group_by(dsp_pedon_id) |>
        dplyr::summarize(depth_wt_bd = mean(bd_hybrid)) |>
        mutate(depth_cat = bd_depth[[i]])
      
    } else {
      out[[i]] <- dice_bd |>
        filter(hrzdep_t >=bd_depth[[i-1]] & hrzdep_t < bd_depth[[i]]) |>
        group_by(dsp_pedon_id) |>
        dplyr::summarize(depth_wt_bd = mean(bd_hybrid)) |>
        mutate(depth_cat = bd_depth[[i]])
    }
}

out_bind <- dplyr::bind_rows(out)

bd_depth <- out_bind |>
  left_join(dplyr::select(project_renamed, project, label, dsp_pedon_id), by="dsp_pedon_id") |>
  rename(layer = depth_cat) |>
  mutate(depth = factor(case_when(layer == "5" ~ "0-5 cm",
                           layer == "10" ~ "5-10 cm",
                           layer == "30" ~ "10-30 cm",
                           layer == "60" ~ "30-60 cm",
                           layer == "100" ~ "60-100 cm"),
                        levels=c("0-5 cm", "5-10 cm", "10-30 cm", "30-60 cm", "60-100 cm")),
         label = factor(label, levels=c("BAU", "SHM", "Ref")),
         project = factor(project, levels=project_plotting_order))

mean_bd_depth <- bd_depth |>
  group_by(project, label, depth) |>
  dplyr::summarize(bd_mean = round(mean(depth_wt_bd, na.rm=TRUE), 2))
```

```{r fig 2 plot bulk density by depth increment}
# Plot bulk density for each project/treatment
ggplot(bd_depth, aes(x=depth, y=depth_wt_bd, fill=label)) +
  geom_point(aes(color=label), pch = 21, position=position_jitterdodge(), size=0.7) +
  geom_boxplot(median.linewidth = 0.3, lwd=0.25, outlier.size=0.6) +
  facet_wrap(~project, scales="free_y", labeller=labeller(project=project_labels_esm)) +
  scale_fill_manual(values=mgmt_pal,
                    breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  scale_color_manual(values=mgmt_pal,
                     breaks=c("BAU", "SHM", "Ref"), 
                     name="Management") +
  labs(x="Depth", y=expression("Bulk density"~(g ~ cm^-3))) +
  theme_katy(base_size=10) +
  theme(axis.text.x=element_text(hjust=1, angle=45))
# ggsave(here("figs", "revision3_figs", "fig2_bd.png"), width=160, height=120, units="mm", dpi=400)
# ggsave(here("figs", "revision3_figs", "fig2.pdf"), width=160, height=120, units="mm", dpi=400)
```

```{r min max bulk density for text}
bd_min_max <- bd_depth |>
  dplyr::summarize(min = min(depth_wt_bd, na.rm=TRUE),
            max = max(depth_wt_bd, na.rm=TRUE))

# or min max mean bulk density
mean_bd_depth <- bd_depth |> 
  group_by(project, label, depth) |> 
  dplyr::summarize(mean = mean(depth_wt_bd, na.rm=TRUE))

mean_bd_depth |> 
  ungroup() |> 
  dplyr::summarize(min_mean = min(mean, na.rm=TRUE),
                   max_mean = max(mean, na.rm=TRUE))
```

## Range of reference masses for different soils:

Full table of reference masses for supplement:
```{r table s2 full ref mass table}
ref_mass_table <- esm |> 
  filter(method_longest != "fd_fd_fd") |> 
  dplyr::select(project, esm_depth, method_longest, soil_mass) |> 
  distinct() |> 
  mutate(soil_mass = round(soil_mass, 0)) |> 
  pivot_wider(names_from = method_longest,
              values_from = soil_mass) |> 
  arrange(project) |> 
  relocate(project, esm_depth, esm2_min_project, esm2_min_treatment, esm2_mean_project, esm2_mean_treatment)

# write_csv(ref_mass_table, here("figs", "revision3_figs", "ref_masses_table.csv"))
```

Summary table for text:

```{r summary table of reference masses for text}
# min and max for each depth
esm |> 
  group_by(depth_std) |> 
  dplyr::summarize(min_mass = round(min(soil_mass), 0),
            max_mass = round(max(soil_mass), 0)) 

esm |> 
  group_by(depth_std, method_longest) |> 
  dplyr::summarize(min_mass = round(min(soil_mass), 0),
            max_mass = round(max(soil_mass), 0)) |> 
  filter(method_longest != "fd_fd_fd")

# mean reference mass for each method/depth
esm |> 
  filter(method_longest != "fd_fd_fd") |> 
  group_by(depth_std, method_longest) |> 
  dplyr::summarize(mean_mass = round(mean(soil_mass, na.rm=TRUE), 0)) |> 
  pivot_wider(names_from = method_longest, values_from = mean_mass)
```

# How does ESM choice influence total calculated SOC stocks?

## Effect of stock calculation method on cumulative SOC stocks

What is the mean SOC stock calculated via each method?

What is the difference between the smallest and largest SOC stocks calculated for a particular pedon using these different methods?

```{r soc stock methods summary}
# summary of projects with the highest and lowest SOC stocks
esm |>
  group_by(depth_std, project, label) |>
  dplyr::summarize(mean = round(mean(soc_cum),1)) |> 
  filter(depth_std == 60) |> 
  ungroup() |> 
  slice(which.max(mean), which.min(mean))

soc_method_summary <- esm |>
  group_by(depth_std, method_longest) |>
  dplyr::summarize(mean = round(mean(soc_cum),1))

soc_method_minmaxdiff <- esm |>
  filter(depth_std=="60") |>
  group_by(project, label, dsp_pedon_id, depth_std) |>
  dplyr::summarize(min_soc_cum = min(soc_cum),
            max_soc_cum = max(soc_cum),
            diff_soc_cum = max_soc_cum - min_soc_cum,
            pct_diff_soc_cum = ((max_soc_cum - min_soc_cum) / min_soc_cum)*100)

mean_diff_project <- soc_method_minmaxdiff |>
  ungroup() |>
  group_by(project) |>
  dplyr::summarize(mean_diff_soc_cum = round(mean(diff_soc_cum), 2),
            min_diff_soc_cum = round(min(diff_soc_cum), 2),
            max_diff_soc_cum = round(max(diff_soc_cum), 2))

mean_diff_project

mean_diff <- soc_method_minmaxdiff |>
  ungroup() |>
  dplyr::summarize(mean_diff_soc_cum = round(mean(diff_soc_cum), 2),
            min_diff_soc_cum = round(min(diff_soc_cum), 2),
            max_diff_soc_cum = round(max(diff_soc_cum), 2),
            max_pct_diff_soc_cum = round(max(pct_diff_soc_cum), 1),
            mean_pct_diff_soc_cum = round(mean(pct_diff_soc_cum), 1))

mean_diff
```

## Effect of stock calculation method on incremental SOC stocks

```{r fig supp2 aov and increment stock calculation method}
aov_soc_inc <- esm |>
  group_by(depth_std) |>
  nest() |>
  mutate(aov = purrr::map(data, ~aov(soc ~ method_longest, data = .x)),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(.x = aov, .y = hsd, ~multcompLetters4(.x, .y, reversed=TRUE)),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[1]]$Letters) |> 
             rownames_to_column() |> 
             rename(method_longest = rowname,
                    letter = 2)
         })
  ) |> 
  dplyr::select(depth_std, letters_df) |> 
  unnest(cols=letters_df)

fig2_inc_letters_position <- esm |>
  group_by(depth_std, method_longest) |>
  dplyr::summarize(max_soc = max(soc)) |>
  left_join(aov_soc_inc, by=c("depth_std", "method_longest"))

ggplot(esm, aes(x=method_longest, y=soc, fill=method_longest)) +
  geom_point(aes(color=method_longest), position = "jitter", alpha=0.8) +
  geom_boxplot(median.linewidth=0.5, lwd=0.25, outlier.shape=NA) +
  geom_text(data=fig2_inc_letters_position, aes(y=max_soc + 30, x=method_longest, label=letter), size=2) +
  labs(x="Calculation method", y=expression("Depth increment SOC stock (Mg ha"^-1*")")) +
  facet_wrap(~depth_std, labeller=labeller(depth_std=inc_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_color_manual(values = method_pal, name="Calculation method") + 
  scale_fill_manual(values = method_pal, name="Calculation method") + 
  theme_katy() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")

# ggsave(here("figs", "revision3_figs", "figsupp2_soc_inc_all.png"), width=120, height=100, units="mm", dpi=400)
```

# How does calculation method influence management sensitivity?

## Calculate means for all treatments, depths, and calculation methods for results text

```{r mean SOC stock in all treatments depths and calc methods}
soc_cum_summary <- esm |> 
  group_by(method_longest, esm_depth, label) |> 
  dplyr::summarize(mean = round(mean(soc_cum), 1)) |> 
  arrange(esm_depth, label, method_longest)

soc_cum_summary |> 
  group_by(esm_depth, label) |> 
  dplyr::summarize(min = min(mean),
            max = max(mean)) 

soc_inc_summary <- esm |> 
  group_by(method_longest, esm_depth, label) |> 
  dplyr::summarize(mean = round(mean(soc), 1)) |> 
    arrange(esm_depth, label, mean)
  
soc_inc_summary |> 
  ungroup() |> 
  group_by(esm_depth) |> 
  dplyr::summarize(min_mean = min(mean),
            max_mean = max(mean))
```

## Management sensitivity of incremental SOC stocks

ANOVA for management sensitivity of incremental SOC stocks using different calculation methods

```{r aov management incremental}
# anova on effect of management on cumulative SOC stocks
mgmt_inc_aov_df <- esm |> 
  group_by(method_longest, depth_std) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(soc ~ label, data=.x)),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, ~multcompLetters4(.x, .y, reversed=TRUE)),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[1]]$Letters) |> 
             rownames_to_column() |> 
             rename(label = rowname,
                    letter = 2)
         }) 
  ) |> 
  dplyr::select(depth_std, method_longest, letters_df) |> 
  unnest(cols=letters_df)

mgmt_inc_aov_letters <- esm |>
  group_by(method_longest, depth_std) |>
  dplyr::summarize(max_soc = max(soc)) |>
  left_join(mgmt_inc_aov_df, by=c("depth_std", "method_longest")) |>
  mutate(label = factor(label, levels=c("BAU", "SHM", "Ref")))
```

```{r fig supp3 soc mgmt incremental aov letters}
ggplot(esm, aes(x=method_longest, y=soc, fill=label)) +
  geom_point(aes(color=label), pch = 21, position=position_jitterdodge(), size=0.7) +
  geom_boxplot(median.linewidth=0.5, lwd=0.25, outlier.shape=NA) +
  scale_x_discrete(labels=method_labels) +
  facet_wrap(~depth_std, scales="free", labeller=labeller(depth_std=inc_depth_labels)) +
  geom_text(data=mgmt_inc_aov_letters, 
            aes(y=max_soc + 25, label=letter, group=label), 
            color="black", position=position_dodge(width=0.8), size=3) +
  labs(x="Calculation method", y=expression("Depth increment SOC (Mg ha"^-1*")")) +
  scale_fill_manual(values=mgmt_pal,
                    breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  scale_color_manual(values=mgmt_pal,
                     breaks=c("BAU", "SHM", "Ref"), 
                     name="Management") +
  theme_katy(base_size=10) +
  theme(axis.text.x=element_text(angle=45, hjust=1))
ggsave(here("figs", "revision3_figs", "fig3_soc_mgmt_inc.png"), width=160, height=130, units="mm", dpi=400)
# ggsave(here("figs", "revision3_figs", "fig3.pdf"), width=120, height=100, units="mm", dpi=400)
```

# Calculation method influence on SOC sequestration

Calculate the mean SOC in the BAU treatment for each project, and then calculate the difference between each Ref/SHM pedon and the BAU mean.

This isn't entirely different from assessing the management sensitivity of each calculation method, as we did above. However, in this analysis we are concerned with the magnitude of the difference in SOC stocks between both Ref and SHM treatments compared to BAU.

```{r calculate difference between BAU and Ref/SHM SOC stocks}
esm_bau <- esm |>
  filter(label=="BAU") |>
  group_by(project, method_long, method_longest, method, ref_stat, ref_data, depth_std) |>
  dplyr::summarize(mean_bau_soil_mass = mean(soil_mass),
            mean_bau_soc = mean(soc),
            mean_bau_soil_mass_cum = mean(soil_mass_cum),
            mean_bau_soc_cum = mean(soc_cum))

esm_diff <- esm |>
  filter(!label=="BAU") |>
  left_join(esm_bau, by=c("project", "method_long", "method_longest", "method", "ref_stat", "ref_data", "depth_std")) |>
  mutate(diff_soil_mass = soil_mass - mean_bau_soil_mass,
         diff_soc = soc - mean_bau_soc,
         diff_soil_mass_cum = soil_mass_cum - mean_bau_soil_mass_cum,
         diff_soc_cum = soc_cum - mean_bau_soc_cum)

# Calculate mean sequestration estimates for each treatment, cumulative depth, and method
esm_diff |> 
  group_by(label, depth_std, method_longest) |> 
  dplyr::summarize(mean_diff = round(mean(diff_soc_cum), 1),
            n = n()) 

# Calculate mean sequestration estimates for each treatment, incremental depth, and method
esm_diff |> 
  group_by(label, depth_std, method_longest) |> 
  dplyr::summarize(mean_dsoc_inc = round(mean(diff_soc), 1)) |> 
  arrange(label, depth_std, mean_dsoc_inc)
```

## Incremental SOC sequestration with different calculation methods

```{r aov calc method dsoc incremental}
dsoc_inc_aov_df <- esm_diff |> 
  group_by(depth_std) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(diff_soc ~ (label+method_longest)^2, data=.x)),
         tidy_aov = purrr::map(aov, broom::tidy),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, ~multcompLetters4(.x, .y, reversed=TRUE)),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[3]]$Letters) |> 
             rownames_to_column() |> 
             rename(letter = 2) |> 
             separate_wider_delim(rowname, ":", names = c("label", "method_longest"))
         }) 
  )

dsoc_inc_aov_tidy <- dsoc_inc_aov_df |> 
  dplyr::select(depth_std, tidy_aov) |> 
  unnest(cols=tidy_aov) |> 
  mutate(comparison = "dsoc_inc") |> 
  relocate(comparison)

dsoc_inc_letters_df <- dsoc_inc_aov_df |> 
  dplyr::select(depth_std, letters_df) |> 
  unnest(cols=letters_df)
# significant differences mostly due to management 

# calculate position for letters on figures
dsoc_inc_quant <- esm_diff |> 
  group_by(depth_std, method_longest, label) |> 
  summarize(quant25 = quantile(diff_soc, probs = 0.25), 
            quant75 = quantile(diff_soc, probs = 0.75)) |> 
  mutate(iqr = quant75-quant25,
         upr = (quant75 + (1.5*iqr))*2.5)

dsoc_inc_letters2 <- dsoc_inc_letters_df |> 
  left_join(dsoc_inc_quant, by=c("method_longest", "depth_std", "label")) |> 
  mutate(label = factor(label, levels = c("SHM", "Ref")))
```

```{r fig3 dsoc incremental}
ggplot(esm_diff, aes(x=method_longest, y=diff_soc, fill=method_longest)) +
  geom_point(aes(color=method_longest), pch=21, position="jitter", alpha=0.8, size=0.8) +
  geom_boxplot(median.linewidth=0.5, lwd=0.25, outlier.shape=NA) +
  geom_abline(intercept=0, slope=0, linetype="dashed", linewidth=0.3) +
  geom_text(data=dsoc_inc_letters2, aes(y=upr, label=letter), size=10/.pt) +
  labs(x="Calculation method", y=expression(Delta*"SOC vs BAU (Mg ha"^-1*")")) +
  facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=inc_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_fill_manual(values = method_pal, name="Calculation method") + 
  scale_color_manual(values = method_pal, name="Calculation method") + 
  theme_katy(base_size=12) +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")
# ggsave(here("figs", "revision3_figs", "fig4_dsoc_inc.png"), width=180, height=160, units="mm", dpi=400)
# ggsave(here("figs", "revision3_figs", "fig4.pdf"), width=180, height=160, units="mm", dpi=400)
```

Plot for each project:

```{r dsoc incremental for each project}
# aov for each project 
dsoc_inc_each_aov <- esm_diff |> 
  group_by(project, depth_std) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(diff_soc ~ (label+method_longest)^2, data=.x)),
         tidy_aov = purrr::map(aov, broom::tidy),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, ~multcompLetters4(.x, .y, reversed=TRUE)),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[3]]$Letters) |> 
             rownames_to_column() |> 
             rename(label.method_longest = rowname,
                    letter = 2) |> 
             separate_wider_delim(label.method_longest, ":", names = c("label", "method_longest"))
         }) 
  )

# anova table
dsoc_inc_each_aov_tidy <- dsoc_inc_each_aov |> 
  dplyr::select(project, depth_std, tidy_aov) |> 
  unnest(cols=tidy_aov) |> 
  mutate(comparison = "dsoc_inc") |> 
  relocate(comparison)

# tukey's HSD letters
dsoc_inc_each_letters <- dsoc_inc_each_aov |> 
  dplyr::select(project, depth_std, letters_df) |> 
  unnest(cols=letters_df) |> 
  group_by(project, depth_std) |> 
  mutate(keep = n_distinct(letter) > 1) |> 
  filter(keep == TRUE)
# very few significant differences

# calculate position for letters on figures
dsoc_inc_each_quant <- esm_diff |> 
  group_by(project, depth_std, method_longest, label) |> 
  summarize(quant25 = quantile(diff_soc, probs = 0.25), 
            quant75 = quantile(diff_soc, probs = 0.75)) |> 
  mutate(iqr = quant75-quant25,
         upr = (quant75 + (1.5*iqr))*1.2)

dsoc_inc_each_letters2 <- dsoc_inc_each_letters |> 
  left_join(dsoc_inc_each_quant, by=c("project","method_longest", "depth_std", "label")) |> 
  mutate(label = factor(label, levels = c("SHM", "Ref")))

purrr::map(.x = projects, 
           .f = ~{
             esm_diff |> 
               filter(project==.x) |> 
               ggplot(aes(x=method_longest, y=diff_soc, fill=method_longest)) +
               geom_point(aes(color=method_longest), pch=21, position="jitter", alpha=0.8, size=0.8) +
               geom_boxplot(median.linewidth=0.5, lwd=0.25, outlier.shape=NA) +
               geom_abline(intercept=0, slope=0, linetype="dashed", linewidth=0.3) +
               geom_text(data=dsoc_inc_each_letters2 |> filter(project == .x), 
                         aes(y=upr, label=letter), size=8/.pt) +
               labs(x="Calculation method", y=expression(Delta*"SOC vs BAU (Mg/ha)"),
                    title=glue::glue("\u0394SOC by calculation method - {project_labels_esm[.x]}")) +
               facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=inc_depth_labels)) +
               scale_x_discrete(labels=method_labels) +
               scale_fill_manual(values = method_pal, name="Calculation method") + 
               scale_color_manual(values = method_pal, name="Calculation method") + 
               theme_classic() +
               theme(axis.text.x=element_text(angle=45, hjust=1),
                     axis.line = element_line(colour = 'black', linewidth = 0.25),
                     strip.background = element_rect(colour = "black", linewidth = 0.35),
                     legend.position="none")
             
             # ggsave(here("figs", "revision3_figs", "proj_dsoc_inc", glue::glue("suppfig_dsoc_inc_", .x, ".png")),
             #                                  width=160, height=140, units="mm", dpi=400)
           })
```

# Calculate error in the delta SOC associated with different methods

First, need to calculate dSOC for ESM1 method. We're considering dSOC calculated with ESM1 to be the "true" value - it was calculated by incrementally going down 1 mm into the soil profile until the reference mass was reached, instead of via cubic spline. So this gets us to the error associated with the cubic spline estimation for each reference mass.

```{r dsoc esm1}
esm1_bau <- esm1 |>
  filter(label=="BAU") |>
  group_by(project, method_long, method_longest, method, ref_stat, ref_data, depth_std) |>
  dplyr::summarize(mean_bau_soil_mass = mean(soil_mass),
            mean_bau_soc = mean(soc),
            mean_bau_soil_mass_cum = mean(soil_mass_cum),
            mean_bau_soc_cum = mean(soc_cum))

esm1_diff <- esm1 |>
  filter(!label=="BAU") |>
  left_join(esm1_bau, by=c("project", "method_long", "method_longest", "method", "ref_stat", "ref_data", "depth_std")) |>
  mutate(diff_soil_mass = soil_mass - mean_bau_soil_mass,
         diff_soc = soc - mean_bau_soc,
         diff_soil_mass_cum = soil_mass_cum - mean_bau_soil_mass_cum,
         diff_soc_cum = soc_cum - mean_bau_soc_cum)
```

```{r visually compare dSOC between esm methods}
esm_diff_all <- bind_rows(esm_diff, esm1_diff) |> 
  unite("stat_data", ref_stat:ref_data, remove = FALSE)

# ESM1 and ESM2 look very similar to each other
ggplot(esm_diff_all, aes(x=stat_data, y=diff_soc, fill=method)) +
  geom_point(aes(color=method), pch=21, position="jitter", alpha=0.8, size=0.8) +
  geom_boxplot(median.linewidth=0.5, lwd=0.25, outlier.shape=NA) +
  geom_abline(intercept=0, slope=0, linetype="dashed", linewidth=0.3) +
  labs(x="Calculation method", y=expression(Delta*"SOC vs BAU (Mg ha"^-1*")")) +
  facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=inc_depth_labels)) +
  scale_fill_manual(values = pnw_palette("Starfish", 3), name="Calculation method") + 
  scale_color_manual(values = pnw_palette("Starfish", 3), name="Calculation method") + 
  theme_katy() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")
```

```{r calculate dsoc error in each depth increment}
# calculate error for ESM methods
esm_inc_error <- esm_diff_all |> 
  dplyr::select(project, stat_data:method, depth_std:diff_soc_cum) |> 
  filter(method!="fd") |> 
  pivot_wider(names_from = method, values_from = mean_bau_soil_mass:diff_soc_cum) |> 
  mutate(inc_delta_soc_error = diff_soc_esm2 - diff_soc_esm1,
         abs_error = abs(inc_delta_soc_error),
         pct_error = abs((diff_soc_esm2 - diff_soc_esm1)/diff_soc_esm1)) |> 
  rename(true_inc_delta_soc = diff_soc_esm1,
         calc_inc_delta_soc = diff_soc_esm2)

# calculate error for FD method
# considering the "true" dSOC for fd method to be dSOC for ESMmean, control
true_dsoc_fd <- esm_diff_all |> 
  dplyr::select(project, stat_data:method, depth_std:diff_soc_cum) |> 
  filter(stat_data == "mean_treatment") |> 
  distinct() |> 
  dplyr::select(project, dsp_pedon_id, layer, depth_std, diff_soc) |> 
  rename(true_inc_delta_soc = diff_soc)
  
esm_inc_error_fd <- esm_diff_all |> 
  dplyr::select(project, stat_data:method, depth_std:diff_soc_cum) |> 
  filter(method == "fd") |> 
  left_join(true_dsoc_fd, by = c("project", "dsp_pedon_id", "layer", "depth_std")) |> 
  mutate(inc_delta_soc_error = diff_soc - true_inc_delta_soc,
         abs_error = abs(inc_delta_soc_error),
         pct_error = abs((diff_soc - true_inc_delta_soc)/true_inc_delta_soc)) |> 
  rename(calc_inc_delta_soc = diff_soc) |> 
  dplyr::select(project:layer, depth_std, calc_inc_delta_soc, true_inc_delta_soc:pct_error)

# join fd error to esm error
esm_inc_error_all <- esm_inc_error |> 
  dplyr::select(project:depth_std, calc_inc_delta_soc, true_inc_delta_soc, inc_delta_soc_error:pct_error) |> 
  bind_rows(esm_inc_error_fd) |> 
  mutate(stat_data = factor(stat_data, levels = method_factor_order2))
```

## Plot at meta-project level

### ANOVA for meta-project

```{r aov delta soc error incremental}
# determine iqr to remove outliers
dsoc_inc_error_quant <- esm_inc_error_all |> 
  group_by(depth_std, stat_data, label) |> 
  summarize(quant25 = quantile(pct_error, probs = 0.25), 
            quant75 = quantile(pct_error, probs = 0.75)) |> 
  mutate(iqr = quant75-quant25,
         upr = quant75 + (1.5*iqr),
         lwr = quant25 - (1.5*iqr),
         upr_label = (quant75 + (1.5*iqr))*2)

# filter outliers out
esm_inc_error_sub <- esm_inc_error_all |> 
  left_join(dplyr::select(dsoc_inc_error_quant, upr, lwr, depth_std, stat_data, label), 
            by=c("label","depth_std", "stat_data")) |> 
  filter(pct_error<=upr, pct_error>=lwr) |> 
  mutate(stat_data = factor(stat_data, levels = method_factor_order2))

# anova on effect of method on percent error
dsoc_inc_error_aov_df <- esm_inc_error_sub |> 
  group_by(depth_std) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(pct_error ~ (label+stat_data)^2, data=.x)),
         tidy_aov = purrr::map(aov, broom::tidy),
         glance_aov = purrr::map(aov, broom::glance),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, multcompLetters4),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[3]]$Letters) |> 
             rownames_to_column() |> 
             rename(letter = 2) |> 
             separate_wider_delim(rowname, ":", names = c("label", "stat_data"))
         }) 
  )

dsoc_inc_error_aov_tidy <- dsoc_inc_error_aov_df |> 
  dplyr::select(depth_std, tidy_aov) |> 
  unnest(cols=tidy_aov) |> 
  mutate(comparison = "dsoc_inc_error") |> 
  relocate(comparison)

# pull out significant depths
dsoc_inc_error_sig <- dsoc_inc_error_aov_tidy |>
  filter(term == "stat_data" | term == "label") |> 
  filter(p.value < 0.05) |> 
  distinct(depth_std) |> 
  pull()

dsoc_inc_error_letters_df <- dsoc_inc_error_aov_df |> 
  dplyr::select(depth_std, letters_df) |> 
  unnest(cols=letters_df) |> 
  group_by(depth_std) |> 
  mutate(keep = n_distinct(letter) > 1) |> 
  filter(keep == TRUE)

dsoc_inc_error_letters_df2 <- dsoc_inc_error_letters_df |> 
  left_join(dsoc_inc_error_quant, by=c("stat_data", "depth_std", "label")) |> 
  mutate(label = factor(label, levels = c("SHM", "Ref")))
```

### Plot at meta project level

```{r fig4 inc dsoc error plot}
ggplot(esm_inc_error_sub, aes(x=stat_data, y=pct_error, fill=stat_data)) +
  geom_point(aes(color=stat_data), pch=21, position="jitter", alpha=0.8, size=0.8) +
  geom_boxplot(median.linewidth = 0.5, lwd=0.25, outliers=FALSE) +
  geom_text(data=dsoc_inc_error_letters_df2, aes(y=upr, label=letter), vjust = -0.4, size=10/.pt) +
  labs(x="Calculation method", y=expression(Delta*"SOC error (%)")) +
  facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=inc_depth_labels)) +
  scale_x_discrete(labels = stat_data_labels) +
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = c(.05, .2))) +
  scale_fill_manual(values = method_pal2, name="Calculation method") +
  scale_color_manual(values = method_pal2, name="Calculation method") +
  theme_katy(base_size=12) +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")
# ggsave(here("figs", "revision3_figs", "fig5_dsoc_inc_error.png"), width=180, height=160, units="mm", dpi=400)
# ggsave(here("figs", "revision3_figs", "fig5.pdf"), width=180, height=160, units="mm", dpi=400)
```

### ANOVA table

```{r dsoc and dsoc inc error anova table}
anova_inc_table <- bind_rows(dsoc_inc_aov_tidy, dsoc_inc_error_aov_tidy) |> 
  dplyr::select(-sumsq, -meansq) |> 
  mutate(statistic = round(statistic, 2),
         p.value = if_else(p.value < 0.001, "<0.001", as.character(round(p.value, 3)))) 
  
anova_inc_table

# write_csv(anova_inc_table, here("figs", "revision3_figs", "dsoc_inc_anova_table.csv"))
```

### Table of means for text

```{r dsoc incremental error table}
esm_inc_error_sub |> 
  group_by(depth_std, stat_data) |> 
  summarize(mean_pct_error = round(mean(pct_error), 3) * 100) |> 
  arrange(depth_std, mean_pct_error)
```

## Incremental error for each project

### ANOVA for each project

```{r dsoc inc error aov for each project}
dsoc_inc_error_each_aov_df <- esm_inc_error_all |> 
  group_by(project, depth_std) |> 
  nest() |> 
  mutate(aov = purrr::map(data, ~aov(abs_error ~ (label+stat_data)^2, data=.x)),
         tidy_aov = purrr::map(aov, broom::tidy),
         glance_aov = purrr::map(aov, broom::glance),
         hsd = purrr::map(aov, TukeyHSD),
         letters = purrr::map2(aov, hsd, multcompLetters4),
         letters_df = purrr::map(letters, .f = ~{
           data.frame(.x[[3]]$Letters) |> 
             rownames_to_column() |> 
             rename(letter = 2) |> 
             separate_wider_delim(rowname, ":", names = c("label", "stat_data"))
         }) 
  )

dsoc_inc_error_each_letters_df <- dsoc_inc_error_each_aov_df |> 
  dplyr::select(project, depth_std, letters_df) |> 
  unnest(cols=letters_df) |> 
  group_by(project, depth_std) |> 
  mutate(keep = n_distinct(letter) > 1) |> 
  filter(keep == TRUE)

# calculate position
dsoc_inc_error_each_quant <- esm_inc_error_all |> 
  group_by(project, depth_std, stat_data, label) |> 
  summarize(quant25 = quantile(abs_error, probs = 0.25), 
            quant75 = quantile(abs_error, probs = 0.75)) |> 
  mutate(iqr = abs(quant75-quant25),
         upr = (quant75 + (1.5*iqr))+1)

dsoc_inc_error_each_letters_df2 <- dsoc_inc_error_each_letters_df |> 
  left_join(dsoc_inc_error_each_quant, by=c("project", "stat_data", "depth_std", "label")) |> 
  mutate(label = factor(label, levels = c("SHM", "Ref")))
```

### Plot for each project

```{r dsoc inc error plot for each project}
purrr::map(.x = projects, 
           .f = ~{
             esm_inc_error_all |> 
               filter(project==.x)  |> 
               ggplot(aes(x=stat_data, y=abs_error, fill=stat_data)) +
               geom_point(aes(color=stat_data), pch=21, position="jitter", alpha=0.8, size=1.3) +
               geom_boxplot(median.linewidth = 0.5, lwd=0.25, outliers=FALSE) +
               geom_abline(intercept=0, slope=0, linetype="dashed", linewidth=0.3) +
               geom_text(data=dsoc_inc_error_each_letters_df2 |> filter(project == .x), 
                         aes(y=upr, label=letter), size=8/.pt) +
               labs(x="Calculation method", y=expression(Delta*"SOC error (Mg/ha)"),
                    title=glue::glue("\u0394SOC error by calculation method - {project_labels_esm[.x]}")) +
               facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=inc_depth_labels)) +
               scale_x_discrete(labels = stat_data_labels) +
               scale_fill_manual(values = method_pal2, name="Calculation method") + 
               scale_color_manual(values = method_pal2, name="Calculation method") + 
               theme_classic() +
               theme(axis.text.x=element_text(angle=45, hjust=1),
                     axis.line = element_line(colour = 'black', linewidth = 0.25),
                     strip.background = element_rect(colour = "black", linewidth = 0.35),
                     legend.position="none")
             
             # ggsave(here("figs", "revision3_figs", "proj_dsoc_error_inc", glue::glue("suppfig_dsoc_error_inc_", .x, ".png")),
             #                                  width=160, height=140, units="mm", dpi=400)
           })
```

# Drivers of dSOC error

## Relationship between bulk density variability and dSOC error

Try calculating the percent difference between an individual pedon's BD in each depth increment and the mean BAU depth increment BD for that project:

```{r calculate pct diff bd}
# first need to calculate pct difference in bulk density between treatments for each depth increment
bd_mean_bau <- bd_depth |> 
  dplyr::select(-depth) |> 
  rename(depth_std = layer) |> 
  group_by(project, label, depth_std) |> 
  summarize(bd_mean_bau = mean(depth_wt_bd)) |> 
  filter(label=="BAU") |> 
  ungroup() |> 
  dplyr::select(-label)

bd_error <- bd_depth |> 
  dplyr::select(-depth) |> 
  rename(depth_std = layer) |> 
  filter(label!="BAU") |> 
  left_join(bd_mean_bau, by=c("project", "depth_std")) |> 
  mutate(bd_diff = depth_wt_bd - bd_mean_bau,
         abs_bd_error = abs(depth_wt_bd - bd_mean_bau),
         pct_error_bd = abs_bd_error / bd_mean_bau)
```

### dSOC error vs BD regression

```{r dsoc error vs bd error regression table}
esm_inc_error_bd_error <- esm_inc_error_sub |> 
  left_join(dplyr::select(bd_error, dsp_pedon_id, depth_std, bd_diff, abs_bd_error, pct_error_bd), 
            by=c("dsp_pedon_id", "depth_std")) |> 
  mutate(abs_true_inc_delta_soc = abs(true_inc_delta_soc))

# all regressions
dsoc_bd_error_lm <- esm_inc_error_bd_error |> 
  group_by(stat_data, depth_std) |> 
  nest() |> 
  mutate(lm = purrr::map(data, ~lm(inc_delta_soc_error ~ bd_diff, data=.x)),
         lm_glance = purrr::map(lm, broom::glance),
         lm_tidy = purrr::map(lm, .f = ~{
           broom::tidy(.x) |> 
             filter(term != "(Intercept)") |> 
             select(estimate)
           })
         )

dsoc_bd_error_lm_table <- dsoc_bd_error_lm |> 
  dplyr::select(stat_data, depth_std, lm_tidy, lm_glance) |> 
  unnest(cols=c(lm_glance, lm_tidy)) |> 
  dplyr::select(stat_data, depth_std, estimate, statistic, df, df.residual, p.value, r.squared) |> 
  mutate(estimate = round(estimate, 2),
         statistic = round(statistic, 2),
         r.squared = round(r.squared, 2),
         p.value = ifelse(p.value < 0.001, "<0.001", as.character(round(p.value, 3)))) |> 
  arrange(stat_data, depth_std) |> 
  relocate(stat_data, depth_std, df, df.residual, statistic, p.value, r.squared)
dsoc_bd_error_lm_table
# write_csv(dsoc_bd_error_lm_table, here("figs", "revision3_figs", "dsoc_bd_regression_table.csv"))
```

```{r sig error dsoc and bd regressions in fd}
fd_error <- esm_inc_error_bd_error |> 
  filter(stat_data == "fd_fd") |> 
  unite("method_depth", c("stat_data", "depth_std"), remove=FALSE)

# first figure out which regressions are significant
fd_reg <- fd_error |>
  group_by(stat_data, depth_std) |>
  nest() |>
  mutate(lm = purrr::map(data, ~lm(inc_delta_soc_error ~ bd_diff, data=.x)),
         lm_tidy = purrr::map(lm, tidy),
         lm_adjr2 = purrr::map(lm, .f = ~{
           glance(.x) |>
             dplyr::select(adj.r.squared)
           }))

fd_sig_reg <- fd_reg |>
  unnest(cols=c(lm_tidy, lm_adjr2)) |>
  filter(term=="bd_diff") |>
  filter(p.value < 0.05) |>
  unite("method_depth", c("stat_data", "depth_std"), remove=FALSE) |>
  mutate(p_print = if_else(p.value < 0.001, "<0.001", paste0("=='", format(round(p.value, 3), nsmall = 3), "'"))) |>
  mutate(reg_lab = paste0("~R^{2} == ", round(adj.r.squared, 2), "*`,`~p", p_print))
```

```{r fig 5 dsoc error vs bd error fixed depth}
ggplot(fd_error, aes(x=bd_diff, y=inc_delta_soc_error)) +
  geom_abline(slope=0, intercept=1, linetype="dashed") +
  geom_vline(xintercept=0, linetype="dashed") +
  geom_point(aes(color=project, shape=label), size = 2) +
  geom_smooth(data=fd_error |>  filter(method_depth %in% fd_sig_reg$method_depth),
              method="lm", se=FALSE, color="black") +
  geom_text(data=fd_sig_reg,
           aes(label = reg_lab, y=50, x=-0.7),
           size = 8/.pt, parse=TRUE) +
  facet_wrap(~depth_std, labeller=labeller(depth_std = inc_depth_labels)) +
  labs(x =expression("Management bulk density difference (g cm"^-3*")"), 
       y = expression("Fixed depth"~Delta*"SOC error (Mg ha"^-1*")")) +
  scale_color_manual(values=pnw_palette("Sunset2", 9), name="Project",
                     labels = project_labels_esm) +
  scale_shape_manual(values = c(16, 17), name="Management") + 
  theme_katy(base_size=10)

# ggsave(here("figs", "revision3_figs", "fig6_fd_dsoc_error_vs_bd_error.png"), width=160, height=120, units="mm", dpi=400)
# ggsave(here("figs", "revision3_figs", "fig6.pdf"), width=160, height=120, units="mm", dpi=400)
```

Also show dSOC error vs BD error for ESM methods:

```{r fig s22 dsoc error vs bd error for esm methods}
# other methods - fig for supplement

# first figure out which regressions are significant
regressions <- esm_inc_error_bd_error |>
  filter(stat_data!="fd_fd") |>
  group_by(stat_data, depth_std) |>
  nest() |>
  mutate(lm = purrr::map(data, ~lm(inc_delta_soc_error ~ bd_diff, data=.x)),
         lm_tidy = purrr::map(lm, tidy),
         lm_adjr2 = purrr::map(lm, .f = ~{
           glance(.x) |>
             dplyr::select(adj.r.squared)
           }))

sig_pos_reg <- regressions |>
  unnest(cols=c(lm_tidy, lm_adjr2)) |>
  filter(term=="bd_diff") |>
  filter(estimate > 0) |> 
  filter(p.value < 0.05) |>
  unite("method_depth", c("stat_data", "depth_std"), remove=FALSE) |>
  mutate(p_print = if_else(p.value < 0.001, "<0.001", paste0("=='", format(round(p.value, 3), nsmall = 3), "'"))) |>
  mutate(reg_lab = paste0("~R^{2} == ", round(adj.r.squared, 2), "*`,`~p", p_print))

error_methods_to_plot <- esm_inc_error_bd_error |> 
  filter(stat_data!="fd_fd") |> 
  unite("method_depth", c("stat_data", "depth_std"), remove=FALSE)

ggplot(error_methods_to_plot, aes(x=bd_diff, y=inc_delta_soc_error)) +
  geom_abline(slope=0, intercept=1, linetype="dashed") +
  geom_vline(xintercept=0, linetype="dashed") +
  geom_point(aes(color=project)) +
  geom_smooth(data=error_methods_to_plot |>  filter(method_depth %in% sig_pos_reg$method_depth),
                           method="lm", color="black", lwd=0.5, se=FALSE) +
  geom_text(data=sig_pos_reg,
           aes(label = reg_lab, y=25, x=-0.4),
           size = 7.5/.pt, parse=TRUE) +
  facet_grid(stat_data~depth_std,
             labeller=labeller(depth_std = inc_depth_labels,
                               stat_data = stat_data_labels2)) +
  labs(x =expression("Management bulk density difference (g cm"^-3*")"), 
       y = expression(Delta*"SOC error (Mg ha"^-1*")")) +
  scale_color_manual(values=pnw_palette("Sunset2", 9), name="Project",
                     labels = project_labels_esm) +  
  theme_katy()
# ggsave(here("figs", "revision3_figs", "figsupp3_all_methods_dsoc_error_vs_bd_error.png"), width=180, height=160, units="mm", dpi=400)
```

Error in FD dSOC compared to ESM(mean) dSOC is significantly correlated with the difference between sample BD and mean BAU bd - in other words, when bulk density changes between treatments, FD results are more biased compared to ESM.

## What's up with the error outliers - relationship between dSOC error and soil mass at each depth?

```{r dsoc error outliers}
error_outliers <- esm_inc_error_all |> 
  mutate(big_error = pct_error > 10) |> 
  filter(big_error == TRUE)

ggplot(error_outliers, aes(x = true_inc_delta_soc, y = abs_error)) +
  geom_point(aes(color = stat_data)) +
  facet_wrap(~depth_std)

# and also find where the error is really low
error_low <- esm_inc_error_all |> 
  mutate(small_error = pct_error < .01) |> 
  filter(small_error = TRUE)
```

Compare linear interpolation to cubic spline interpolation

```{r fig6 plot actual mass vs depth}
# for a single profile, plot linear vs cubic spline interpolation

linear <- horizon_data |> 
  mutate(hrzdepth = hrzdep_b - hrzdep_t,
         mass = hrzdepth * bd_hybrid * 100,
         soc_stock = soc_fill * bd_hybrid * hrzdepth) |> 
  group_by(dsp_pedon_id) |> 
  mutate(cum_mass = cumsum(mass),
         cum_soc = cumsum(soc_stock))

# one example with large error - CeH1-3 from the NC State project, which had 1400% error in the 30-60 ESM depth

# filter data
ceh13_pts <- linear |> 
  select(dsp_pedon_id, cum_mass, cum_soc) |> 
  filter(dsp_pedon_id == "CeH1-3") 

ceh13_line <- ceh13_pts |> 
  ungroup() |> 
  add_row(dsp_pedon_id = "CeH1-3", cum_mass = 0, cum_soc = 0, .before = 1)

ceh13_esm <- esm |> 
  filter(dsp_pedon_id == "CeH1-3")

# make plots
min_error_plot2 <- ggplot() +
  geom_point(data = ceh13_esm |>  filter(method_longest == "esm2_min_project"), 
             aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
             size = 2) +
  geom_line(data = ceh13_esm |>  filter(method_longest == "esm2_min_project"), 
            aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
            linewidth = 1) +
  geom_point(data = ceh13_pts, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
             size = 2, pch = 8) +
  geom_line(data = ceh13_line, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
            linewidth = 0.8, linetype = "dashed") +
    ylim(c(19000,0)) +
  labs(x = expression("Cumulative SOC (Mg ha"^-1*")"),
       y = expression("Cumulative soil mass (Mg ha"^-1*")"),
       title = "SHM",
       subtitle = expression("ESM"["project, min"])) +
  scale_color_manual(values = c("esm2_min_project" = "#4a9152",
                                "CeH1-3" = "black"), 
                     labels = c("esm2_min_project" = "Cubic spline interpolation",
                                "CeH1-3" = "Linear interpolation"),
                     name = "") + 
  theme_classic()

mean_error_plot2 <- ggplot() +
  geom_point(data = ceh13_esm |>  filter(method_longest == "esm2_mean_treatment"), 
             aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
             size = 2) +
  geom_line(data = ceh13_esm |>  filter(method_longest == "esm2_mean_treatment"), 
            aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
            linewidth = 1) +
  geom_point(data = ceh13_pts, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
             size = 2, pch = 8) +
  geom_line(data = ceh13_line, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
            linewidth = 0.8, linetype = "dashed") +
    ylim(c(19000,0)) +
  labs(x = expression("Cumulative SOC (Mg ha"^-1*")"),
       y = expression("Cumulative soil mass (Mg ha"^-1*")"),
       subtitle = expression("ESM"["control, mean"])) +
  scale_color_manual(values = c("esm2_mean_treatment" = "#4a9152",
                                "CeH1-3" = "black"), 
                     labels = c("esm2_mean_treatment" = "Cubic spline interpolation",
                                "CeH1-3" = "Linear interpolation"),
                     name = "") + 
  theme_classic()

max_error_plot2 <- ggplot() +
  geom_point(data = ceh13_esm |>  filter(method_longest == "esm2_max_project"), 
             aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
             size = 2) +
  geom_line(data = ceh13_esm |>  filter(method_longest == "esm2_max_project"), 
            aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
            linewidth = 1) +
  geom_point(data = ceh13_pts, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
             size = 2, pch = 8) +
  geom_line(data = ceh13_line, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
            linewidth = 0.8, linetype = "dashed") +
  ylim(c(19000,0)) +
  labs(x = expression("Cumulative SOC (Mg ha"^-1*")"),
       y = expression("Cumulative soil mass (Mg ha"^-1*")"),
       subtitle = expression("ESM"["project, max"])) +
  scale_color_manual(values = c("esm2_max_project" = "#4a9152",
                                "CeH1-3" = "black"), 
                     labels = c("esm2_max_project" = "Cubic spline interpolation",
                                "CeH1-3" = "Linear interpolation"),
                     name = "") + 
  theme_classic()

# another row - pick a soil profile in Ref
# Ref - KeN1-1
ken1_pts <- linear |> 
  select(dsp_pedon_id, cum_mass, cum_soc) |> 
  filter(dsp_pedon_id == "KeN1-1")

ken1_line <- ken1_pts |> 
  ungroup() |> 
  add_row(dsp_pedon_id = "KeN1-1", cum_mass = 0, cum_soc = 0, .before = 1)

ken1_esm <- esm |> 
  filter(dsp_pedon_id == "KeN1-1")

# make plots
min_error_plot3 <- ggplot() +
  geom_point(data = ken1_esm |>  filter(method_longest == "esm2_min_project"), 
             aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
             size = 2) +
  geom_line(data = ken1_esm |>  filter(method_longest == "esm2_min_project"), 
            aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
            linewidth = 1) +
  geom_point(data = ken1_pts, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
             size = 2, pch = 8) +
  geom_line(data = ken1_line, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
            linewidth = 0.8, linetype = "dashed") +
    ylim(c(12000,0)) +
  labs(x = expression("Cumulative SOC (Mg ha"^-1*")"),
       y = expression("Cumulative soil mass (Mg ha"^-1*")"),
       title = "Ref",
       subtitle = expression("ESM"["project, min"])) +
  scale_color_manual(values = c("esm2_min_project" = "#4a9152",
                                "KeN1-1" = "black"), 
                     labels = c("esm2_min_project" = "Cubic spline interpolation",
                                "KeN1-1" = "Linear interpolation"),
                     name = "") + 
  theme_classic()

mean_error_plot3 <- ggplot() +
  geom_point(data = ken1_esm |>  filter(method_longest == "esm2_mean_treatment"), 
             aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
             size = 2) +
  geom_line(data = ken1_esm |>  filter(method_longest == "esm2_mean_treatment"), 
            aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
            linewidth = 1) +
  geom_point(data = ken1_pts, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
             size = 2, pch = 8) +
  geom_line(data = ken1_line, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
            linewidth = 0.8, linetype = "dashed") +
    ylim(c(12000,0)) +
  labs(x = expression("Cumulative SOC (Mg ha"^-1*")"),
       y = expression("Cumulative soil mass (Mg ha"^-1*")"),
       subtitle = expression("ESM"["control, mean"])) +
  scale_color_manual(values = c("esm2_mean_treatment" = "#4a9152",
                                "KeN1-1" = "black"), 
                     labels = c("esm2_mean_treatment" = "Cubic spline interpolation",
                                "KeN1-1" = "Linear interpolation"),
                     name = "") + 
  theme_classic()

max_error_plot3 <- ggplot() +
  geom_point(data = ken1_esm |>  filter(method_longest == "esm2_max_project"), 
             aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
             size = 2) +
  geom_line(data = ken1_esm |>  filter(method_longest == "esm2_max_project"), 
            aes(x = soc_cum, y = soil_mass_cum, color = method_longest),
            linewidth = 1) +
  geom_point(data = ken1_pts, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
             size = 2, pch = 8) +
  geom_line(data = ken1_line, aes(y = cum_mass, x = cum_soc, color = dsp_pedon_id), 
            linewidth = 0.8, linetype = "dashed") +
  ylim(c(12000,0)) +
  labs(x = expression("Cumulative SOC (Mg ha"^-1*")"),
       y = expression("Cumulative soil mass (Mg ha"^-1*")"),
       subtitle = expression("ESM"["project, max"])) +
  scale_color_manual(values = c("esm2_max_project" = "#4a9152",
                                "KeN1-1" = "black"), 
                     labels = c("esm2_max_project" = "Cubic spline interpolation",
                                "KeN1-1" = "Linear interpolation"),
                     name = "") + 
  theme_classic()

# put plots together
min_error_plot2 + mean_error_plot2 + max_error_plot2 +
min_error_plot3 + mean_error_plot3 + max_error_plot3 +
  plot_layout(axes = "collect", guides = "collect")

# ggsave(here("figs", "revision3_figs", "interpolation_profiles_alt.png"), width = 190, height = 140, units = "mm")
```

Okay, this is actually really helpful! The thicker, purple line shows the results of linear interpolation, aka "classical" ESM in which the soil profile is discretized and 1 mm layers are summed in order to reach the reference mass. Looks here like most of the "errors" associated with the cubic spline interpolation occur in the big jump in SOC mass in the deeper depths...honestly the cubic spline is more likely to be correct.

Also plot the soil profiles:

```{r get soil horizon colors from database}
# need to connect to database and get colors from data description

# Connect to database
db <- dbConnect(RSQLite::SQLite(), here("..", "dsp4sh_shiny", "data", "dsp4sh6.db"))

pedons <- c("CeH1-3", "KeN1-1")

# sql query
sql <- sprintf("SELECT
    desg.dsp_pedon_id, desg.dsp_sample_id, desg.hzdesg, desg.hrzdep_t, desg.hrzdep_b,
    descr.color_moist_hue, descr.color_moist_value, descr.color_moist_chroma
  FROM layerdesignation AS desg
  JOIN layerdescription AS descr ON descr.dsp_sample_id = desg.dsp_sample_id
  WHERE desg.dsp_pedon_id in %s
  ", format_SQL_in_statement(pedons)
)

pedon_data <- dbGetQuery(db, sql) |> 
  clean_names()

hex_colors <- munsell2rgb(pedon_data$color_moist_hue, pedon_data$color_moist_value, pedon_data$color_moist_chroma)

pedons_w_color <- pedon_data |> 
  mutate(soil_color = hex_colors)

# done with connection
dbDisconnect(db)
```

```{r plot soil profiles}
cecil <- pedons_w_color |> 
  filter(dsp_pedon_id == "CeH1-3")

depths(cecil) <- dsp_pedon_id ~ hrzdep_t + hrzdep_b
cecil$soil_series <- "Cecil"

plotSPC(cecil, name = "hzdesg", label = "soil_series", cex.names = 1, cex.id = 1.2)

keith <- pedons_w_color |> 
  filter(dsp_pedon_id == "KeN1-1")

depths(keith) <- dsp_pedon_id ~ hrzdep_t + hrzdep_b
keith$soil_series <- "Keith"

plotSPC(keith, name = "hzdesg", label = "soil_series", cex.names = 1, cex.id = 1.2)
```


# Check that ESM calculations worked and make sense

## Plot mean cumulative SOC vs depth, and cumulative SOC vs cumulative soil mass

These plots are just to check that the ESM calculations worked and nothing looks too crazy - for example, wild fluctuations in SOC content with depth, impossible numbers, etc.

Cumulative mass vs depth:

```{r cumulative mass vs depth}
ggplot(esm, aes(x=depth_std, y=soil_mass_cum, linetype=method_longest, color=project)) +
  stat_summary(geom="line", fun="mean", linewidth=1.3) +
  coord_flip() +
  xlim(c(60,0)) +
  labs(x="ESM Depth (cm)", y="Cumulative mass (Mg/ha)", 
       title="Cumulative mass with depth") +
  scale_color_manual(values=pnw_palette("Sunset2", 9), name="Project",
                     labels = project_labels_esm) +  
  theme_classic()
```

Cumulative SOC vs depth:

```{r cumulative soc vs depth}
purrr::map(.x = projects, 
           .f = ~{
             esm |>
               filter(project==.x) |>
               ggplot(aes(x=depth_std, y=soc_cum, color=method_longest)) +
               stat_summary(geom="line", fun="mean", linewidth=1.3) +
               coord_flip() +
               xlim(c(60,0)) +
               facet_grid(~label, scales="free") +
               labs(x="ESM Depth (cm)", y="Cumulative SOC (Mg/ha)", 
                    title=glue::glue("Cumulative SOC with depth - ", .x)) +
               scale_color_manual(values = method_pal, name="Calculation method", labels=method_labels) + 
               theme_classic()
           })
```

These aren't the easiest to interpret, but I think that everything looks good here. It doesn't look like SOC estimates take a crazy left turn at any depth for any of the methods, and they are all in relative agreement with each other. The plot of cumulative mass vs cumulative SOC shows that using the minimum reference mass for ESM results in a lower cumulative soil mass overall, but that's no surprise. Fixed depth results look pretty identical to ESM mean.
 
## How does ESM choice influence soil mass in each depth increment?

This is another way to check that nothing is getting wonky with the ESM calculations - making sure that the soil mass in each depth increment makes sense for each project.
 
 Plot soil mass in each depth increment:
 
```{r soil mass in each depth increment} 
# Soil mass in each depth increment
purrr::map(.x = projects, 
           .f = ~{
             esm |>
               filter(project==.x) |>
               ggplot(aes(x=esm_depth, y=soil_mass, fill=method_longest)) +
               stat_summary(geom="bar", fun=mean, position="dodge2", color="black") +
               stat_summary(geom="errorbar", fun.data = mean_cl_normal, color="black",
                            width=0.4, position=position_dodge(width=0.9)) +
               coord_flip() +
               scale_x_discrete(limits=rev) +
               facet_wrap(~label) +
               labs(x="ESM Depth", y="Depth Increment Soil Mass (Mg/ha)", 
                    title=glue::glue("Depth increment soil mass by calculation method - ", .x)) +
               scale_fill_manual(values = method_pal,
                                 name="Calculation method",
                                 labels=method_labels) + 
               theme_classic()
           })
```

Plot cumulative soil mass in each depth increment:

```{r cumulative soil mass in each depth increment}
# Cumulative soil mass in each depth increment
purrr::map(.x = projects, 
           .f = ~{
             esm |>
               filter(project==.x) |>
               ggplot(aes(x=factor(depth_std), y=soil_mass_cum, fill=method_longest)) +
               stat_summary(geom="bar", fun=mean, position="dodge2", color="black") +
               stat_summary(geom="errorbar", fun.data=mean_cl_normal, color="black",
                            width=0.4, position=position_dodge(width = 0.9)) +
               coord_flip() +
               scale_x_discrete(limits=rev) +
               facet_wrap(~label) +
               labs(x="ESM Depth (cm)", y="Cumulative Soil Mass (Mg/ha)", 
                    title=glue::glue("Cumulative soil mass by calculation method - ", .x)) +
               scale_fill_manual(values = method_pal, 
                                 name="Calculation method",
                                 labels=method_labels) + 
               theme_classic()
           })
```

These plots are a check to see how using ESM methods influence the mass of soil in a given depth increment. They show that, unsurprisingly, fixed depth and ESM(max) calculations result in the greatest soil mass in each depth increment. Fixed depth soil masses are most comparable to ESM(mean) masses.