---
title: "esm_manuscript_revision"
author: "Katy Dynarski"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages
library(here)
library(janitor)
library(paletteer)
library(aqp)
library(broom)
library(broom.mixed)
library(lme4)
library(lmerTest)
library(multcomp)
library(ggeffects)
library(flextable)
library(patchwork)
library(multcomp)
library(multcompView)
library(tidyverse)

# Data
# Vector for method order for consistent plotting
method_factor_order <- c("esm2_min_project", "esm2_min_treatment", "esm2_mean_project", "esm2_mean_treatment", "fd_fd_fd")
# Vector for project order for consistent plotting
project_plotting_order <- c("UConn", "Illinois", "KansasState", "UnivOfMinnesota", "NCState", 
                            "OregonStateJory", "OregonStateWoodburn","TexasA&MPt-1", "WashingtonState")

esm <- read.csv(here("data_processed", "esm_standard_depths_revision.csv")) %>%
  filter(project!="UTRGV") %>% # don't include UTRGV for this analysis - bulk density is from SSURGO
  filter(method!="esm1") %>% # not including ESM classic in this analysis - just compare cubic spline
  unite("method_long", c("method", "ref_stat"), sep="_", remove = FALSE) %>%
  unite("method_longest", c("method", "ref_stat", "ref_data"), sep="_", remove = FALSE) %>%
  mutate(depth_std = case_when(layer==1 ~ 5,
                               layer==2 ~ 10,
                               layer==3 ~ 30,
                               layer==4 ~ 60),
         esm_depth = factor(esm_depth, levels=c("0-5 cm", "5-10 cm", "10-30 cm", "30-60 cm")),
         label = factor(label, levels=c("BAU", "SHM", "Ref")),
         project = factor(project, levels=project_plotting_order),
         method_longest = factor(method_longest, levels=method_factor_order))
  

horizon_data <- read.csv(here("data_processed", "soc_horizon_esm_input.csv"))
horizon_text_data <- read.csv(here("data_processed", "soc_horizon_clim_meta.csv"))
project <- read.csv(here("data_processed", "project_data_meta.csv"))

# List of projects
projects <- esm %>% 
  distinct(project) %>% 
  mutate(project = as.character(project)) %>%
  pull()

# Make list of summary functions - useful for generating summary tables
sd_cv <- list(
  sd = ~round(sd(.x, na.rm = TRUE), 2),
  cv = ~round(((sd(.x, na.rm=TRUE) / mean(.x, na.rm=TRUE))* 100), 2)
)


# depth layer labels
depth_labels1 <- c(
  "5" = "0-5 cm",
  "10" = "5-10 cm",
  "30" = "10-30 cm",
  "60" = "30-60 cm"
)

depth_labels2 <- c(
  "5" = "ESM Depth: 0-5 cm",
  "10" = "ESM Depth: 5-10 cm",
  "30" = "ESM Depth: 10-30 cm",
  "60" = "ESM Depth: 30-60 cm"
)

cum_depth_labels <- c(
  "5" = "ESM Depth: 0-5 cm",
  "10" = "ESM Depth: 0-10 cm",
  "30" = "ESM Depth: 0-30 cm",
  "60" = "ESM Depth: 0-60 cm"
)
  
method_labels <- c(
  "esm2_min_project" = expression("ESM"["project, min"]),      
  "esm2_mean_project" = expression("ESM"["project, mean"]), 
  "esm2_min_treatment"= expression("ESM"["treatment, min"]),   
  "esm2_mean_treatment" = expression("ESM"["treatment, mean"]),
  "fd_fd_fd" = "Fixed depth"
)

method_labels2 <- c(
  "esm2_min_project" = "ESM(project, min)",      
  "esm2_mean_project" = "ESM(project, mean)", 
  "esm2_min_treatment"= "ESM(treatment, min)",   
  "esm2_mean_treatment" = "ESM(treatment, mean)",
  "fd_fd_fd" = "Fixed depth"
)

project_labels_esm <- c("UnivOfMinnesota" = "U. Minnesota",
                    "WashingtonState" = "Washington State",
                    "TexasA&MPt-1" = "Texas A&M",
                    "NCState" = "North Carolina State",
                    "Illinois" = "U. Illinois",
                    "OregonStateJory" = "Oregon State - Jory",
                    "OregonStateWoodburn" = "Oregon State - Woodburn",
                    "KansasState" = "Kansas State",
                    "UConn" = "U. Connecticut")

project_labels_esm_df <- data.frame(project_labels_esm) %>%
  rownames_to_column()

mgmt_labels <- c("Ref" = "Perennial reference",
                 "SHM" = "Soil health agriculture")

theme_katy <- function(base_size=8) {
  theme_classic(base_size=base_size) %+replace%
    theme(# Legend
      legend.title=element_text(size=rel(1)), legend.text=element_text(size=rel(.8)),
      #axes                      
      axis.text=element_text(size=rel(.8)),
      axis.title=element_text(size=rel(1)))
}
```

# Overview

## Calculation method comparison

I calculated SOC stocks in 0-5 cm, 5-10 cm, 10-30 cm, and 30-60 cm depth increments via fixed depth and ESM methods. For the ESM calculations, I tested four different reference mass options:

* Minimum soil mass in an individual DSP4SH project (ESM project, min)
* Mean soil mass in an individual DSP4SH project (ESM project, mean)
* Minimum soil mass in the reference treatment in an individual DSP4SH project (ESM treatment, min)
* Mean soil mass in the reference treatment in an individual DSP4SH project (ESM treatment, mean)

I wanted to test just using data from the reference treatment to calculate the reference mass because a number of studies calculated their reference mass as either the mean or minimum soil mass in the treatment expected to have the lowest bulk density.

I also re-calculated bulk density after reading through the SimpleESM documentation again. We did not originally account for coarse fragment content, so I re-calculated bulk density to correct for coarse fragment content. It looks like this resolved some of the weird numbers we were getting for soils high in coarse frags (e.g. the Oregon State project).

# Climate and soils data for each project 

```{r climate means for each project}
# need to rename OSU projects in "project" dataframe and calculate mean MAT/MAP for all projects
project_renamed <- project %>%
  mutate(project = case_when(
    project=="OregonState" & soil=="Jory" ~ "OregonStateJory",
    project=="OregonState" & soil=="Woodburn" ~ "OregonStateWoodburn", 
    .default = project))

project_avg_climate <- project_renamed %>%
  group_by(project) %>%
  dplyr::summarize(avg_mat = round(mean(mat, na.rm=TRUE), 1),
         avg_ppt = round(mean(map, na.rm=TRUE), 0)) %>%
  arrange(factor(project, levels=project_plotting_order))

# write_csv(project_avg_climate, here("figs", "revision_figs", "tablesupp1_site_clim.csv"))
```

# Number of pedons and sites

```{r how many pedons and sites}

# how many projects
n_distinct(esm$project)

# how many pedons
n_distinct(esm$dsp_pedon_id)

# how many sites
# have to use some regex to extract the site name from the pedon ID since I'm missing the needed column
sites <- esm %>%
  select(project, label, dsp_pedon_id) %>%
  distinct(project, label, dsp_pedon_id) %>%
  mutate(dsp_site_id = str_remove(dsp_pedon_id, "-\\d*\\w*$")) %>%
  group_by(project, label, dsp_site_id)

# how many pedons per site?
sites %>%
  dplyr::summarize(n_pedons_per_site = n())

# how many total sites  
n_distinct(sites$dsp_site_id)

# how many total pedons
n_distinct(sites$dsp_pedon_id)

# how many total pedons per treatment
esm %>%
  select(project, label, dsp_pedon_id) %>%
  distinct(project, label, dsp_pedon_id) %>%
  group_by(label) %>%
  dplyr::summarize(n_pedons_per_treatment = n())

# table for supplement with number of sites and pedons per treatment in each project
site_table <- sites %>%
  distinct(project, label, dsp_site_id) %>%
  group_by(project, label) %>%
  dplyr::summarize(n_sites_per_treatment = n())

pedon_table <- sites %>%
  ungroup() %>%
  group_by(project, label) %>%
  dplyr::summarize(n_pedons_per_trt = n())

site_pedon_table <- site_table %>%
  left_join(pedon_table, by=c("project", "label")) %>%
  arrange(factor(project, levels=project_plotting_order), label)

# write_csv(site_pedon_table, here("figs", "revision_figs", "tablesupp2_site_pedon_numbers.csv"))
```

# Range in bulk density and reference masses

Management can influence soil bulk density, which results in different masses of soil within the same depth increment between treatments. We can check this by plotting soil bulk density under the different management treatments in each project.

Plot bulk density:

```{r calculate bulk density by depth increment, include=FALSE}
bd <- horizon_data %>%
  filter(!project=="UTRGV") %>%
  select(project, label, dsp_pedon_id, dsp_sample_id, hrzdep_t, hrzdep_b, bd_hybrid)

# promote to SPC and dice into 1-cm increments
spc <- bd
aqp::depths(spc) <- dsp_pedon_id ~ hrzdep_t + hrzdep_b

dice <- aqp::dice(spc, fm=0:99 ~ bd_hybrid + project + label)
dice_bd <- horizons(dice)

bd_depth <- c(5, 10, 30, 60, 100)

# Initialize output vector
out <- vector("list", length(bd_depth))

for (i in seq_along(bd_depth)) {
    if (i == 1) {
      out[[i]] <- dice_bd %>%
        filter(hrzdep_t < bd_depth[[i]]) %>%
        group_by(dsp_pedon_id) %>%
        dplyr::summarize(depth_wt_bd = mean(bd_hybrid)) %>%
        mutate(depth_cat = bd_depth[[i]])
      
    } else {
      out[[i]] <- dice_bd %>%
        filter(hrzdep_t >=bd_depth[[i-1]] & hrzdep_t < bd_depth[[i]]) %>%
        group_by(dsp_pedon_id) %>%
        dplyr::summarize(depth_wt_bd = mean(bd_hybrid)) %>%
        mutate(depth_cat = bd_depth[[i]])
    }
}

out_bind <- dplyr::bind_rows(out)

bd_depth <- out_bind %>%
  left_join(select(project_renamed, project, label, dsp_pedon_id), by="dsp_pedon_id") %>%
  rename(layer = depth_cat) %>%
  mutate(depth = factor(case_when(layer == "5" ~ "0-5 cm",
                           layer == "10" ~ "5-10 cm",
                           layer == "30" ~ "10-30 cm",
                           layer == "60" ~ "30-60 cm",
                           layer == "100" ~ "60-100 cm"),
                        levels=c("0-5 cm", "5-10 cm", "10-30 cm", "30-60 cm", "60-100 cm")),
         label = factor(label, levels=c("BAU", "SHM", "Ref")),
         project = factor(project, levels=project_plotting_order))

mean_bd_depth <- bd_depth %>%
  group_by(project, label) %>%
  dplyr::summarize(bd_mean = round(mean(depth_wt_bd, na.rm=TRUE), 2))
```

```{r fig s2 plot bulk density by depth increment}
# Plot bulk density for each project/treatment
ggplot(bd_depth, aes(x=depth, y=depth_wt_bd, fill=label)) +
  geom_boxplot(fatten=1.5, lwd=0.3) +
  facet_wrap(~project, scales="free_y", labeller=labeller(project=project_labels_esm)) +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  labs(x="Depth", y=expression("Bulk density"~(g ~ cm^-3))) +
  theme_classic() +
  theme(axis.text.x=element_text(hjust=1, angle=45))

# ggsave(here("figs", "revision_figs", "figsupp2_bulk_density.png"), width=200, height=160, units="mm", dpi=400)
```

```{r min max bulk density for text}
bd_min_max <- bd_depth %>%
  dplyr::summarize(min = min(depth_wt_bd, na.rm=TRUE),
            max = max(depth_wt_bd, na.rm=TRUE))
```

```{r table 1 calculate within treatment CV vs full project CV}
bd_soc <- esm %>%
  left_join(dplyr::select(bd_depth, -depth), by=c("dsp_pedon_id", "project", "depth_std" = "layer", "label"))

treatment_bd_cv <- bd_soc %>%
  filter(method=="fd", project!="Illinois", project!="UConn") %>%
  group_by(project, label, depth_std) %>%
  dplyr::summarize(mean_bd = mean(depth_wt_bd),
            sd_bd = sd(depth_wt_bd)) %>%
  mutate(cv_bd_treatment = sd_bd/mean_bd)

overall_treatment_cv <- treatment_bd_cv %>%
  ungroup() %>%
  group_by(depth_std, label) %>%
  dplyr::summarize(treatment_mean_cv = mean(cv_bd_treatment, na.rm=TRUE))

project_bd_cv <- bd_soc %>%
  filter(method=="fd", project!="Illinois", project!="UConn") %>%
  group_by(project, depth_std) %>%
  dplyr::summarize(mean_bd = mean(depth_wt_bd),
            sd_bd = sd(depth_wt_bd)) %>%
  mutate(cv_bd_project = sd_bd/mean_bd)

overall_project_cv <- project_bd_cv %>%
  ungroup() %>%
  group_by(depth_std) %>%
  dplyr::summarize(project_mean_cv = mean(cv_bd_project, na.rm=TRUE))

bd_cv_table <- overall_treatment_cv %>%
  left_join(overall_project_cv, by="depth_std") %>%
  mutate(across(treatment_mean_cv:project_mean_cv, ~round(.x, 2))) %>%
  arrange(depth_std, label)

# write_csv(bd_cv_table, here("figs", "revision_figs", "table1_bd_cv.csv"))
```

## Range of reference masses for different soils
```{r fig 2 range of reference masses in each ESM depth}
mean_ref_mass <- esm %>%
  group_by(depth_std) %>%
  dplyr::summarize(mean_mass = mean(soil_mass),
            mean_mass_cum = mean(soil_mass_cum))

cust_labeller <- function(x) paste0("ESM Depth: ", x)

ggplot(esm %>% filter(method!="fd"), aes(x=project, y=soil_mass, fill=method)) +
  geom_boxplot() +
  facet_wrap(~esm_depth, scales="free", labeller=as_labeller(cust_labeller)) +
  scale_fill_paletteer_d("nationalparkcolors::Arches",
                         name="Calculation method") + 
  labs(x="DSP4SH project", y="Reference soil mass (Mg/ha)") +
  scale_x_discrete(labels=project_labels_esm) +
  theme_classic() +
    theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")

# ggsave(here("figs", "revision_figs", "fig2_soil_mass.png"), width=180, height=160, units="mm", dpi=400)
# ggsave(here("figs", "revision_figs", "fig2.pdf"), width=180, height=160, units="mm", dpi=400)
```

```{r table of reference masses for text}
ref_mass_table <- esm %>%
  group_by(depth_std) %>%
  dplyr::summarize(min_mass = min(soil_mass),
            max_mass = max(soil_mass))
```

# How does ESM choice influence total calculated SOC stocks?

## Effect of stock calculation method on cumulative SOC stocks

What is the mean SOC stock calculated via each method?

What is the difference between the smallest and largest SOC stocks calculated for a particular pedon using these different methods?

```{r soc stock methods summary}
soc_method_summary <- esm %>%
  group_by(depth_std, method_longest) %>%
  dplyr::summarize(mean = round(mean(soc_cum),1))

soc_method_minmaxdiff <- esm %>%
  filter(depth_std=="60") %>%
  group_by(project, label, dsp_pedon_id, depth_std) %>%
  dplyr::summarize(min_soc_cum = min(soc_cum),
            max_soc_cum = max(soc_cum),
            diff_soc_cum = max_soc_cum - min_soc_cum,
            pct_diff_soc_cum = ((max_soc_cum - min_soc_cum) / min_soc_cum)*100)

mean_diff_project <- soc_method_minmaxdiff %>%
  ungroup() %>%
  group_by(project) %>%
  dplyr::summarize(mean_diff_soc_cum = round(mean(diff_soc_cum), 2),
            min_diff_soc_cum = round(min(diff_soc_cum), 2),
            max_diff_soc_cum = round(max(diff_soc_cum), 2))

flextable(mean_diff_project)

mean_diff <- soc_method_minmaxdiff %>%
  ungroup() %>%
  dplyr::summarize(mean_diff_soc_cum = round(mean(diff_soc_cum), 2),
            min_diff_soc_cum = round(min(diff_soc_cum), 2),
            max_diff_soc_cum = round(max(diff_soc_cum), 2),
            max_pct_diff_soc_cum = round(max(pct_diff_soc_cum), 1),
            mean_pct_diff_soc_cum = round(mean(pct_diff_soc_cum), 1))

flextable(mean_diff)
```

Run linear mixed effects model:

```{r lmer calculation method}
soc_method_lmer <- esm %>%
  group_by(depth_std) %>%
  nest() %>%
  mutate(lmer = purrr::map(data, ~lmer(soc_cum ~ method_longest + (1|project/label), data = .x)),
  drop1 = purrr::map(data, .f = ~{
    drop1(lmer(soc_cum ~ method_longest + (1|project/label), REML=FALSE, data = .x))
    }),
  letters = purrr::map(lmer, .f = ~{
    glht <- glht(.x, linfct = mcp(method_longest = "Tukey"))
    cld <- cld(glht)
    tidy(cld)
    }),
  tidy_lmer = purrr::map(lmer, broom.mixed::tidy),
  tidy_drop = purrr::map(drop1, broom.mixed::tidy),
  predict = purrr::map(lmer, ~ggpredict(.x, terms = c("method_longest")))) 

# Extract table of F-values and p-values for each depth
method_drop_table <- soc_method_lmer %>%
  select(depth_std, tidy_drop) %>%
  unnest(cols=c(tidy_drop), names_sep="_") %>%
  mutate(sig = case_when(tidy_drop_p.value < 0.05 ~ "significant",
                         tidy_drop_p.value > 0.05 ~ "not_significant")) %>%
  select(depth_std, tidy_drop_statistic, tidy_drop_p.value, sig) %>%
  distinct() %>%
  mutate(across(where(is.numeric), ~round(.x, 2)),
         tidy_drop_p.value = ifelse(tidy_drop_p.value<0.001, "<0.001", tidy_drop_p.value))

flextable(method_drop_table)

# Extract significantly different management groups for each method
method_cld_table <- soc_method_lmer %>%
  select(depth_std, letters) %>%
  unnest(cols=c(letters))

flextable(method_cld_table)
```

These plots show that (unsurprisingly), using different calculation methods results in significantly different SOC stocks, especially in surface soil horizons. Using ESM min methods results in the smallest stocks (of course), while fixed depth tends to be more comparable to ESM mean estimates.

```{r fig 3 plot actual cumulative SOC stocks at each depth}
fig2_letters_position <- esm %>%
  group_by(depth_std, method_longest) %>%
  dplyr::summarize(max_soc_cum = max(soc_cum)) %>%
  left_join(method_cld_table, by=c("depth_std", "method_longest"))

ggplot(esm, aes(x=method_longest, y=soc_cum, fill=method_longest)) +
  geom_point(aes(color=method_longest), position = "jitter", alpha=0.8) +
  geom_boxplot(outlier.shape=NA) +
  geom_text(data=fig2_letters_position, aes(y=max_soc_cum + 30, x=method_longest, label=letters), size=3) +
  labs(x="Calculation method", y="Cumulative SOC stock (Mg/ha)") +
  facet_wrap(~depth_std, labeller=labeller(depth_std=cum_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_color_paletteer_d("nationalparkcolors::Arches",
                          name="Calculation method") + 
  scale_fill_paletteer_d("nationalparkcolors::Arches",
                          name="Calculation method") + 
  theme_classic() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")

# ggsave(here("figs", "revision_figs", "fig3_soc_cum_all.png"), width=160, height=120, units="mm", dpi=400)
# ggsave(here("figs", "revision_figs", "fig3.pdf"), width=160, height=120, units="mm", dpi=400)
```

# How does calculation method influence management sensitivity?

## Calculate means for all treatments, depths, and calculation methods for results text

```{r mean SOC stock in all treatments depths and calc methods}
soc_cum_summary <- esm %>%
  group_by(method_longest, esm_depth, label) %>%
  dplyr::summarize(mean = round(mean(soc_cum), 1)) %>%
  arrange(esm_depth, label, method_longest)

soc_cum_summary2 <- soc_cum_summary %>%
  group_by(esm_depth, label) %>%
  dplyr::summarize(min = min(mean),
            max = max(mean))
```

## Management sensitivity of cumulative SOC stocks

We can test the overall sensitivity to management by running a mixed effects linear model that accounts for depth and project:

```{r lmer management}
soc_mgmt_lmer <- esm %>%
  group_by(method_longest, depth_std) %>%
  nest() %>%
  mutate(lmer = purrr::map(data, ~lmer(soc_cum ~ label + (1|project), data = .x)),
  drop1 = purrr::map(data, .f = ~{
    drop1(lmer(soc_cum ~ label + (1|project), REML=FALSE, data = .x))
    }),
  letters = purrr::map(lmer, .f = ~{
    glht <- glht(.x, linfct = mcp(label = "Tukey"))
    cld <- cld(glht)
    tidy(cld)
    }),
  tidy_lmer = purrr::map(lmer, broom.mixed::tidy),
  tidy_drop = purrr::map(drop1, broom.mixed::tidy),
  predict = purrr::map(lmer, ~ggpredict(.x, terms = c("label")))) 

# Extract table of F-values and p-values for each method
mgmt_drop_table <- soc_mgmt_lmer %>%
  select(method_longest, depth_std, tidy_drop) %>%
  unnest(cols=c(tidy_drop), names_sep="_") %>%
    select(method_longest, tidy_drop_statistic, tidy_drop_p.value) %>%
  mutate(tidy_drop_p.value = ifelse(tidy_drop_p.value < 0.05,  "<0.001", as.character(tidy_drop_p.value)),
         tidy_drop_statistic = round(tidy_drop_statistic, 2),
         method_longest=factor(method_longest, levels=method_factor_order)) %>%
  arrange(method_longest, depth_std) %>%
  relocate(method_longest, .before=depth_std)

# Extract significantly different management groups for each method
mgmt_cld_table <- soc_mgmt_lmer %>%
  select(method_longest, depth_std, letters) %>%
  unnest(cols=c(letters))

tablesupp2 <- mgmt_drop_table %>%
  left_join(mgmt_cld_table, by=c("method_longest", "depth_std")) %>%
  mutate(label = factor(label, levels=c("BAU", "SHM", "Ref")),
         method_longest=factor(method_longest, levels=method_factor_order)) %>%
  arrange(method_longest, depth_std, label)

flextable(tablesupp2)

# write_csv(tablesupp2, here("figs", "revision_figs", "tablesupp3_mgmt.csv"))
```

When accounting for differences in depth and project, all calculation methods result in SOC stocks that are highest in Ref condition, and lower in both SHM and BAU conditions.

Plot actual differences due to management:

```{r fig 4 plot actual cumulative SOC stocks under different management}
mgmt_letters_df <- esm %>%
  group_by(method_longest, depth_std) %>%
  dplyr::summarize(max_soc_cum = max(soc_cum)) %>%
  left_join(mgmt_cld_table, by=c("depth_std", "method_longest")) %>%
  mutate(label = factor(label, levels=c("BAU", "SHM", "Ref")))

ggplot(esm, aes(x=method_longest, y=soc_cum, fill=label)) +
  geom_point(aes(color=label), pch = 21, position=position_jitterdodge(), alpha=0.8) +
  geom_boxplot(outlier.shape=NA) +
  scale_x_discrete(labels=method_labels) +
  facet_wrap(~depth_std, scales="free", labeller=labeller(depth_std=cum_depth_labels)) +
  geom_text(data=mgmt_letters_df, 
            aes(y=max_soc_cum + 25, label=letters, group=label), 
            color="black", position=position_dodge(width=0.8), size=3) +
  labs(x="Calculation method", y="Cumulative SOC (Mg/ha)") +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                    breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  scale_color_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                     name="Management") +
  theme_classic() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

# ggsave(here("figs", "revision_figs", "fig4_soc_mgmt.png"), width=180, height=160, units="mm", dpi=400)
# ggsave(here("figs", "revision_figs", "fig4.pdf"), width=180, height=160, units="mm", dpi=400)
```

# Calculation method influence on SOC sequestration

Calculate the mean SOC in the BAU treatment for each project, and then calculate the difference between each Ref/SHM pedon and the BAU mean.

This isn't entirely different from assessing the management sensitivity of each calculation method, as we did above. However, in this analysis we are concerned with the magnitude of the difference in SOC stocks between both Ref and SHM treatments compared to BAU.

```{r calculate difference between BAU and Ref/SHM SOC stocks}
esm_bau <- esm %>%
  filter(label=="BAU") %>%
  group_by(project, method_long, method_longest, method, ref_stat, ref_data, depth_std) %>%
  dplyr::summarize(mean_bau_soil_mass = mean(soil_mass),
            mean_bau_soc = mean(soc),
            mean_bau_soil_mass_cum = mean(soil_mass_cum),
            mean_bau_soc_cum = mean(soc_cum))

esm_diff <- esm %>%
  filter(!label=="BAU") %>%
  left_join(esm_bau, by=c("project", "method_long", "method_longest", "method", "ref_stat", "ref_data", "depth_std")) %>%
  mutate(diff_soil_mass = soil_mass - mean_bau_soil_mass,
         diff_soc = soc - mean_bau_soc,
         diff_soil_mass_cum = soil_mass_cum - mean_bau_soil_mass_cum,
         diff_soc_cum = soc_cum - mean_bau_soc_cum)

# Calculate mean sequestration estimates for each treatment, depth, and method
esm_diff_summary <- esm_diff %>%
  group_by(label, depth_std, method_longest) %>%
  dplyr::summarize(mean_diff = round(mean(diff_soc_cum), 1),
            n = n())
flextable(esm_diff_summary)
```

## Cumulative SOC sequestration with different calculation methods

```{r lmer for influence of calculation method on differences in cumulative soc stocks}

diff_cum_method_lmer <- esm_diff %>%
  group_by(depth_std, label) %>%
  nest() %>%
  mutate(lmer = purrr::map(data, ~lmer(diff_soc_cum ~ method_longest + (1|project), data = .x)),
  drop1 = purrr::map(data, .f = ~{
    drop1(lmer(diff_soc_cum ~ method_longest + (1|project), REML=FALSE, data = .x))
    }),
  letters = purrr::map(lmer, .f = ~{
    glht <- glht(.x, linfct = mcp(method_longest = "Tukey"))
    cld <- cld(glht)
    tidy(cld)
    }),
  tidy_lmer = purrr::map(lmer, broom.mixed::tidy),
  tidy_drop = purrr::map(drop1, broom.mixed::tidy),
  predict = purrr::map(lmer, ~ggpredict(.x, terms = c("method_longest")))) 

# Extract table of F-values and p-values for each depth
diff_cum_method_drop_table <- diff_cum_method_lmer %>%
  unnest(cols=c(tidy_drop), names_sep="_") %>%
  mutate(sig = case_when(tidy_drop_p.value < 0.05 ~ "significant",
                         tidy_drop_p.value > 0.05 ~ "not_significant")) %>%
  select(depth_std, label, tidy_drop_statistic, tidy_drop_p.value, sig) %>%
  distinct() %>%
  mutate(across(where(is.numeric), ~round(.x, 2)))

flextable(diff_cum_method_drop_table)

# Extract significantly different management groups for each method
diff_cum_method_cld_table <- diff_cum_method_lmer %>%
  select(depth_std, label, letters) %>%
  unnest(cols=c(letters))

flextable(diff_cum_method_cld_table)
```

```{r fig 5 plot actual differences in cumulative SOC stocks with different methods}
diff_cum_letters_position <- esm_diff %>%
  group_by(method_longest, depth_std, label) %>%
  dplyr::summarize(max_diff_soc_cum = max(diff_soc_cum)) %>%
  left_join(diff_cum_method_cld_table, by=c("depth_std", "label", "method_longest"))

ggplot(esm_diff, aes(x=method_longest, y=diff_soc_cum, fill=method_longest)) +
  geom_point(aes(color=method_longest), pch=21, position="jitter", alpha=0.8) +
  geom_boxplot(outlier.shape=NA) +
  geom_abline(intercept=0, slope=0, linetype="dashed") +
  geom_text(data=diff_cum_letters_position, aes(y=max_diff_soc_cum + 25, label=letters), size=3) +
  labs(x="Calculation method", y="SOC stock difference vs BAU (Mg/ha)") +
  facet_grid(depth_std~label, scales="free", labeller=labeller(depth_std=cum_depth_labels)) +
  scale_x_discrete(labels=method_labels) +
  scale_fill_paletteer_d("nationalparkcolors::Arches",
                         name="Calculation method") + 
  scale_color_paletteer_d("nationalparkcolors::Arches",
                         name="Calculation method") + 
  theme_classic() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position="none")

# ggsave(here("figs", "revision_figs", "fig5_soc_diff.png"), width=180, height=170, units="mm", dpi=400)
# ggsave(here("figs", "revision_figs", "fig5.pdf"), width=180, height=170, units="mm", dpi=400)
```

# Check that ESM calculations worked and make sense

## Plot mean cumulative SOC vs depth, and cumulative SOC vs cumulative soil mass

These plots are just to check that the ESM calculations worked and nothing looks too crazy - for example, wild fluctuations in SOC content with depth, impossible numbers, etc.

Cumulative SOC vs depth:

```{r cumulative soc vs depth}
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=depth_std, y=soc_cum, color=method_longest)) +
               stat_summary(geom="line", fun="mean", linewidth=1.3) +
               coord_flip() +
               xlim(c(60,0)) +
               facet_grid(~label, scales="free") +
               labs(x="ESM Depth (cm)", y="Cumulative SOC (Mg/ha)", 
                    title=glue::glue("Cumulative SOC with depth - ", .x)) +
               scale_color_paletteer_d("nationalparkcolors::Arches",
                                       name="Calculation method", labels=method_labels) + 
               theme_classic()
           })
```

Cumulative SOC vs cumulative soil mass:

```{r cumulative soc vs cumulative soil mass}
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=soil_mass_cum, y=soc_cum, color=label, group=dsp_pedon_id)) +
               geom_smooth(stat="identity") +
               coord_flip() +
               scale_x_reverse() +
               facet_wrap(~method_longest, labeller=labeller(method_longest=method_labels2)) +
               labs(x="Cumulative soil mass (Mg/ha)", y="Cumulative SOC (Mg/ha)", 
                    title=glue::glue("Cumulative SOC vs cumulative soil mass - ", .x)) +
               scale_color_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                                  breaks=c("BAU", "SHM", "Ref"), 
                                  name="Management") +
               theme_classic()
           })
```
 
These aren't the easiest to interpret, but I think that everything looks good here. It doesn't look like SOC estimates take a crazy left turn at any depth for any of the methods, and they are all in relative agreement with each other. The plot of cumulative mass vs cumulative SOC shows that using the minimum reference mass for ESM results in a lower cumulative soil mass overall, but that's no surprise. Fixed depth results look pretty identical to ESM mean.
 
## How does ESM choice influence soil mass in each depth increment?

This is another way to check that nothing is getting wonky with the ESM calculations - making sure that the soil mass in each depth increment makes sense for each project.
 
 Plot soil mass in each depth increment:
 
```{r soil mass in each depth increment} 
# Soil mass in each depth increment
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=esm_depth, y=soil_mass, fill=method_longest)) +
               stat_summary(geom="bar", fun=mean, position="dodge2", color="black") +
               stat_summary(geom="errorbar", fun.data = mean_cl_normal, color="black",
                            width=0.4, position=position_dodge(width=0.9)) +
               coord_flip() +
               scale_x_discrete(limits=rev) +
               facet_wrap(~label) +
               labs(x="ESM Depth", y="Depth Increment Soil Mass (Mg/ha)", 
                    title=glue::glue("Depth increment soil mass by calculation method - ", .x)) +
               scale_fill_paletteer_d("nationalparkcolors::Arches",
                                      name="Calculation method",
                                      labels=method_labels) + 
               theme_classic()
           })
```

Plot cumulative soil mass in each depth increment:

```{r cumulative soil mass in each depth increment}
# Cumulative soil mass in each depth increment
purrr::map(.x = projects, 
           .f = ~{
             esm %>%
               filter(project==.x) %>%
               ggplot(aes(x=factor(depth_std), y=soil_mass_cum, fill=method_longest)) +
               stat_summary(geom="bar", fun=mean, position="dodge2", color="black") +
               stat_summary(geom="errorbar", fun.data=mean_cl_normal, color="black",
                            width=0.4, position=position_dodge(width = 0.9)) +
               coord_flip() +
               scale_x_discrete(limits=rev) +
               facet_wrap(~label) +
               labs(x="ESM Depth (cm)", y="Cumulative Soil Mass (Mg/ha)", 
                    title=glue::glue("Cumulative soil mass by calculation method - ", .x)) +
               scale_fill_paletteer_d("nationalparkcolors::Arches",
                                      name="Calculation method",
                                      labels=method_labels) + 
               theme_classic()
           })
```

These plots are a check to see how using ESM methods influence the mass of soil in a given depth increment. They show that, unsurprisingly, fixed depth calculations result in the greatest soil mass in each depth increment. Fixed depth soil masses are most comparable to ESM(mean) masses.